### 2018-03-26 Paper.tex
Initial commit; adding files to new repository

\documentclass[12pt]{article}
\usepackage{}
%%%%%%%%%%%%%%%%%%%%%%%%
% PREAMBLE
\usepackage{times}
\usepackage{setspace}
\linespread{1.25}
\usepackage[letterpaper, margin=1in]{geometry}
\addtolength{\topmargin}{-0.5in}

\begin{document}
\title{Constructive Solid Geometry for Interactive Environments}
\author{Nicholas Czaban}
\date{\today}
\maketitle

\section{Constructive Solid Geometry}
\subsection{General Concepts}
Constructive Solid Geometry (CSG) is a method for creating more complicated shapes by performing boolean operations on simpler shapes. The available boolean operations include unions, intersections, and differences; the types of shapes vary from one implementation to another, but generally include spheres, rectangular prisms, and cylinders.
\subsubsection{Real-Time Rendering and Image-Based CSG}
One substantial category in CSG is the difference between Real-Time Rendering of CSG objects and Image-Based CSG objects. Real-time rendering poses a number of issues, particularly for shapes that are interactively modified during runtime. While real-time rendering can be accomplished by passing the shape as a mathematically calculated model, this solution is more applicable to static CSG shapes and does not work well for interactive modifications\cite{open_csg}. The alternative solution, image-based rendering algorithms, avoid most of the complex boundary calculations of constructive solid geometry by storing only an image of the shape in the frame buffer\cite{open_csg}. 
\subsection{Use in Computer Graphics}
\subsubsection{Interactive Modification}
\subsubsection{Ray Tracing}
\subsection{CSG Algorithms}
\section{Software}
\subsection{Display Program}
\subsection{Asteroids}
\subsubsection{CSG Usage}
The shapes are individually created using OpenSCAD, a program for computer assisted design. The shapes are exported as .stl files, which can be read using the reader function located in stlReader.cpp.
\subsubsection{OpenGL Implementation}
\section{Conclusion}

\newpage
\begin{thebibliography}{10}
\bibitem{cs_dictionary}
  Andrew Butterfield and Gerard Ngondi,
  \textit{A Dictionary of Computer Science},
  Oxford University Press,
  7th edition,
  2016

\bibitem{hardware_csg}
  Kirsch, Florian and J\"{u}rgen D\"{o}llner. ``Rendering Techniques for Hardware-Accelerated Image-Based CSG.'' {\it Journal of WSCG}, vol. 12, no. 2, 2004, pp. 221-228.

\bibitem{open_csg}
  Kirsch, Florian and J\"{u}rgen D\"{o}llner. ``OpenCSG: A Library for Image-Based CSG Rendering.'' {\it Proceedings of the FREENIX/ Open Source Track, 2005 USENIX Annual Technical Conference}. 2005.

\bibitem{advanced_opengl}
  McReynolds, Tom and David Blythe. {\it Advanced Graphic Programming Using OpenGL}. Morgan Kaufmann, San Francisco, 2005.

\bibitem{geoinformation}
  Ooms, Kristien. ``A 3D inclusion test on large dataset.'' {\it Developments in 3D Geo-Information Sciences}. Edited by Philippe De Maeyer and Tijs Neutens. Springer, New York, 2010.

\bibitem{raytrace}
  Wilt, Nicholas. {\it Object-Oriented Ray Tracing in C++}. John Wiley \& Sons, New York, 1994, pp 207-227.

\bibitem{interactive_csg}
  Ulyanov, D., D. Bogolepov, and V. Turlapov. ``Interactive Visualization of Constructive Solid Geometry Scenes on Graphic Processors.'' {\it Programming and Computer Software}, vol. 43, no. 4, 2017, pp. 258-267.
\end{thebibliography}
\end{document}

### 2018-03-26 asteroids.cpp

@@ -1,0 +1,252 @@
+#ifdef __APPLE__
+#  include <OpenGL/gl.h>
+#  include <OpenGL/glu.h>
+#  include <GLUT/glut.h>
+#else
+#  include "GL/freeglut.h"
+#endif
+
+#include "stlReader.cpp"
+
+// Name: Nicholas Czaban
+// Assignment: Final Project 
+// Title: CSG Asteroid Game
+// Course: CS 200 
+// Semester: Spring year
+// Instructor: D. Byrnes
+// Date: 2/21/18
+// Sources consulted: any books consulted in writing the program
+// Program description: A simple game where the user navigates a spaceship through an asteroid.
+// Known bugs: 
+// Creativity: 
+// Instructions: Compile and run on the command line.
+
+GLint window_width = 700;
+GLint window_height = 700;
+static int moving = 0, startx = 0, starty = 0;
+static GLfloat angle = 0;
+static GLfloat angle2 = 0;
+static GLfloat xpos = 0;
+static GLfloat ypos = 0;
+static GLfloat zpos = 0;
+static GLfloat speed = 0.2;
+
+GLfloat light_specular[] = {1, 1, 1, 1.0};
+GLfloat light_position[] = {0.0, -2.0, -10.0, 0.0};
+
+std::vector<Facet> asteroids = readFile("pipes.stl");
+std::vector<Facet> spaceship = readFile("spaceship.stl");
+
+bool upKey, leftKey, rightKey, downKey;
+
+void setup_scene(){
+  // Draw the spaceship
+  glPushMatrix();
+  glRotatef(-1*angle2, 1.0, 0.0, 0.0);
+  glRotatef(-1*angle, 0.0, 1.0, 0.0);
+
+  glColor3f(0.4, 0.4, 0.4);
+  glTranslatef(xpos, ypos, 1);
+  if(leftKey) glRotatef(15, 0, 0, 1);
+  if(rightKey) glRotatef(-15, 0, 0, 1);
+  if(upKey) glRotatef(-15, 1, 0, 0);
+  if(downKey) glRotatef(15, 1, 0, 0);
+
+  /****
+	   The stl files generated by OpenSCAD use the right-hand rule for its coordinate system.
[…]

### 2018-03-26 display.cpp

@@ -1,0 +1,136 @@
+#ifdef __APPLE__
+#  include <OpenGL/gl.h>
+#  include <OpenGL/glu.h>
+#  include <GLUT/glut.h>
+#else
+#  include "GL/freeglut.h"
+#endif
+
+#include "stlReader.cpp"
+
+// Name: Nicholas Czaban
+// Assignment: Final Project 
+// Title: CSG Display Program
+// Course: CS 200 
+// Semester: Spring year
+// Instructor: D. Byrnes
+// Date: 2/19/18
+// Sources consulted: any books consulted in writing the program
+// Program description: Using the function in stlReader.cpp, create a model of a CSG object with OpenGL's TRIANGLES primitive.
+// Known bugs: Program does not check to see if the file is of type STL before running. On occasion, some triangles on the shape will not be drawn correctly; this issue is resolved on repeat executions.
+// Creativity: 
+// Instructions: Compile and run on the command line. Type the name of the STL file as the first command line option when running. Click and move the mouse to rotate the object in space. Press 'q' to exit the program.
+
+GLint window_width = 500;
+GLint window_height = 500;
+static int moving = 0, startx = 0, starty = 0;
+static GLfloat angle = 0;
+static GLfloat angle2 = 0;
+std::string filename = "";
+
+void display(){
+  glClear(GL_COLOR_BUFFER_BIT | GL_DEPTH_BUFFER_BIT);
+  std::vector<Facet> newShape = readFile(filename);
+
+  GLfloat light_specular[] = {1.0, 1.0, 1.0, 1.0};
+  GLfloat light_position[] = {0.0, 2.0, -10.0, 0.0};
+
+  GLfloat mat_shininess[] = {50.0};
+  GLfloat mat_specular[] = {1.0, 0.0, 0.0, 1.0};
+  
+  glLightfv(GL_LIGHT0, GL_SPECULAR, light_specular);
+  glLightfv(GL_LIGHT0, GL_POSITION, light_position);
+
+  glPushMatrix();
+  glRotatef(-1*angle2, 1.0, 0.0, 0.0);
+  glRotatef(-1*angle, 0.0, 1.0, 0.0);
[…]

### 2018-03-26 game.cpp

@@ -1,0 +1,164 @@
+#ifdef __APPLE__
+#  include <OpenGL/gl.h>
+#  include <OpenGL/glu.h>
+#  include <GLUT/glut.h>
+#else
+#  include "GL/freeglut.h"
+#endif
+
+#include "stlReader.cpp"
+
+// Name: Nicholas Czaban
+// Assignment: Final Project 
+// Title: CSG Asteroid Game
+// Course: CS 200 
+// Semester: Spring year
+// Instructor: D. Byrnes
+// Date: 2/21/18
+// Sources consulted: any books consulted in writing the program
+// Program description: A simple game where the user navigates a spaceship through an asteroid.
+// Known bugs: 
+// Creativity: 
+// Instructions: Compile and run on the command line.
+
+GLint window_width = 700;
+GLint window_height = 700;
+static int moving = 0, startx = 0, starty = 0;
+static GLfloat angle = 0;
+static GLfloat angle2 = 0;
+static GLfloat xpos = 0;
+static GLfloat ypos = 0;
+static GLfloat zpos = 0;
+static GLfloat speed = 0.2;
+
+std::vector<Facet> asteroids = readFile("pipes.stl");
+std::vector<Facet> spaceship = readFile("spaceship.stl");
+
+bool upKey, leftKey, rightKey, downKey;
+
+void test_draw(){
+  gluLookAt(0,0,0,
+			0,0,1,
+			0,0,1);
+  glBegin(GL_TRIANGLES);
+  glVertex3f(-1,-1,1);
+  glVertex3f(0,1,1);
+  glVertex3f(1,-1,1);
+  glEnd();
+}
+
+void display (void) {
+	glClear(GL_COLOR_BUFFER_BIT | GL_DEPTH_BUFFER_BIT);
+	glMatrixMode(GL_MODELVIEW);
+	glLoadIdentity();
+	test_draw();
+	glutSwapBuffers();
+}
+
+void init(){
+  glClearColor(0.1, 0.1, 0.1, 1.0);
+  glMatrixMode(GL_PROJECTION);
+  glLoadIdentity();
+  gluPerspective(85, window_width/window_height, 1, 25);
+  
+  glShadeModel(GL_SMOOTH);
+  glEnable(GL_CULL_FACE);
+  glEnable(GL_DEPTH_TEST);
+  glEnable(GL_LIGHTING);
[…]

### 2018-03-26 stlReader.cpp

@@ -1,0 +1,96 @@
+// Name: Nicholas
+// Assignment: Final Project 
+// Title: STL File Reader
+// Course: CS 200 
+// Semester: Spring year
+// Instructor: D. Byrnes
+// Date: 2/19/18
+// Program description: Reads STL files created by OpenSCAD and creates a vector of Facets
+// Known Bugs: The reader can only handle STL facets with three vertices and specific headings as provided by OpenSCAD. It is unclear whether or not every STL file generated by OpenSCAD or other programs will use the same format
+// Creativity:
+// Instructions: With this file in your working directory, use a #include "stlReader.cpp" in your program and call the functions below
+
+#include <fstream>
+#include <iostream>
+#include <vector>
+#include <string>
+#include <string.h>
+#include <stdlib.h>
+
+using std::endl;
+using std::cout;
+using std::string;
+
+struct Facet{
+  double normal[3]; // The x,y,z values of the normal vector
+  double vertices[3][3]; // The x,y,z values of the three vertices for each Facet. [][i] covers the x,y,z values, while [j][] handles the points
+};
+
+std::ostream& operator <<(std::ostream &o, const Facet& f){
+  o << endl;
+  o << "Normal Vector: x=" << f.normal[0] << " y=" << f.normal[1] << " z=" << f.normal[2] << endl;
+  for(int i = 0; i < 3; i++){
+	o << "Vertex " << i+1 << ": x=" << f.vertices[i][0] << " y=" << f.vertices[i][1] << " z=" << f.vertices[i][2] << endl;
+  }
+  o << endl;
+}
+
+std::vector<Facet> readFile(std::string filename){
+  std::vector<Facet> facets;
+  std::ifstream file(filename.c_str());
+  string line;
+  if(file.is_open()){
+	Facet tmpFacet;
+	int vertexNum = 0;
+	while(std::getline(file, line)){
+	  int pos=0, found;
+	  char str[line.length()];
+	  for(int i = 0; i < line.length(); i++){
+		str[i] = line[i];
+	  }
+	  char* token;
+	  try{
+		token = strtok(str, " ");
+		string sToken(token);
+		if(sToken == "facet"){
+		  token = strtok(NULL, " ");
+		  token = strtok(NULL, " ");
+		  tmpFacet.normal[0] = atof(token);
+		  token = strtok(NULL, " ");
[…]

### 2018-03-27 game.cpp
Fixing gluLookAt function in game.cpp

@@ -25,11 +25,14 @@ GLint window_width = 700;
 GLint window_height = 700;
 static int moving = 0, startx = 0, starty = 0;
 static GLfloat angle = 0;
-static GLfloat angle2 = 0;
+static GLfloat angle2 = 180;
 static GLfloat xpos = 0;
 static GLfloat ypos = 0;
 static GLfloat zpos = 0;
 static GLfloat speed = 0.2;
+
+GLfloat light_specular[] = {1, 1, 1, 1.0};
+GLfloat light_position[] = {0.0, -2.0, -10.0, 0.0};
 
 std::vector<Facet> asteroids = readFile("pipes.stl");
 std::vector<Facet> spaceship = readFile("spaceship.stl");
@@ -37,21 +40,72 @@ std::vector<Facet> spaceship = readFile("spaceship.stl");
 bool upKey, leftKey, rightKey, downKey;
 
 void test_draw(){
-  gluLookAt(0,0,0,
-			0,0,1,
-			0,0,1);
-  glBegin(GL_TRIANGLES);
-  glVertex3f(-1,-1,1);
-  glVertex3f(0,1,1);
-  glVertex3f(1,-1,1);
+  gluLookAt(0,0,-5,
+			0,0,0,
+			0,1,0);
+  glColor3f(1,1,1);
+  glBegin(GL_POLYGON);
+  glVertex3f(-1,-1,0);
+  glVertex3f(0,1,0);
+  glVertex3f(1,-1,0);
   glEnd();
+}
+
+void setup_scene(){
+  // Draw the spaceship
+  glPushMatrix();
+  glRotatef(-1*angle2, 1.0, 0.0, 0.0);
+  glRotatef(-1*angle, 0.0, 1.0, 0.0);
+
+  glColor3f(0.4, 0.4, 0.4);
+  glTranslatef(xpos, ypos, 1);
+  if(leftKey) glRotatef(15, 0, 0, 1);
+  if(rightKey) glRotatef(-15, 0, 0, 1);
+  if(upKey) glRotatef(-15, 1, 0, 0);
+  if(downKey) glRotatef(15, 1, 0, 0);
+
+  /****
+	   The stl files generated by OpenSCAD use the right-hand rule for its coordinate system.
+	   Thus, in order to correctly render the shapes, the z-coordinates are multiplied by -1.
+   ****/
+  
[…]

### 2018-03-31 Paper.tex
Wrote more of the paper

@@
 \usepackage{times}
+\usepackage{courier}
 \usepackage{setspace}
@@
 \subsection{General Concepts}
-Constructive Solid Geometry (CSG) is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
-The available boolean operations include unions, intersections, and differences; the types of shapes vary from one implementation to another, but generally include spheres, rectangular prisms, and cylinders.
+\indent Constructive Solid Geometry (CSG) is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
+The available boolean operations include unions, intersections, and differences; the types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations)\cite{cs_dictionary}.
 \subsubsection{Real-Time Rendering and Image-Based CSG}
-One substantial category in CSG is the difference between Real-Time Rendering of CSG objects and Image-Based CSG objects.
+\indent One substantial category in CSG is the difference between Real-Time Rendering of CSG objects and Image-Based CSG objects.
 Real-time rendering poses a number of issues, particularly for shapes that are interactively modified during runtime.
@@
 \subsection{CSG Algorithms}
+\indent Given the rendering complexity of constructive solid geometry, several techniques have been developed to improve the speed and quality of CSG primitives.
+In Nicholas Wilt's implementation of CSG in the Object Oriented Ray Tracing (OORT) class library, complicated calculations can be short-circuited by creating optimized CSG trees where simpler objects are the left children of each boolean operation.
+In OORT, ray tracing calculations between two objects on a CSG tree always test against the left object first.
+Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE\cite{raytrace}.
 \section{Software}
@@
 \subsubsection{CSG Usage}
-The shapes are individually created using OpenSCAD, a program for computer assisted design. The shapes are exported as .stl files, which can be read using the reader function located in stlReader.cpp.
+The shapes are individually created using OpenSCAD, a program for computer assisted design.
+The shapes are exported as .stl files, which can be read using the reader function located in \texttt{stlReader.cpp}.
 \subsubsection{OpenGL Implementation}

### 2018-03-31 Paper.tex
Adding code snippets to paper

@@
 \usepackage{setspace}
+\usepackage{textcomp}
+\usepackage{listings}
 \linespread{1.25}
@@
 \addtolength{\topmargin}{-0.5in}
+
+\lstset{language=C++, upquote=true}
+
+\lstdefinelanguage{scad}
+{
+  morekeywords={true, false},
+  sensitive=true,
+  commentstyle=\it,
+  morecomment=[l][//],
+  morestring=[b]''
+  breaklines=true,
+  showstringspaces=false,
+}
 
@@
 The available boolean operations include unions, intersections, and differences; the types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations)\cite{cs_dictionary}.
+This technique provides a simpler way to create three-dimensional shapes without the need to define each vertex
 \subsubsection{Real-Time Rendering and Image-Based CSG}
@@
 \subsection{Display Program}
+
 \subsection{Asteroids}
+\indent 
 \subsubsection{CSG Usage}
-The shapes are individually created using OpenSCAD, a program for computer assisted design.
+\indent The shapes are individually created using OpenSCAD, a program for computer assisted design.
 The shapes are exported as .stl files, which can be read using the reader function located in \texttt{stlReader.cpp}.
+The code within a standard .scad file will look similar to this:
+\lstset{language=scad}
+\begin{lstlisting}
+  // A 6-sided die with rounded corners and circular faces
+  intersection() {
+    cube(15, center=true);
+    sphere(r=10, \$fn=50);
+  }
+
+\end{lstlisting}
 \subsubsection{OpenGL Implementation}
 \section{Conclusion}
-
+\newpage
+\section{Appendix}
+\subsection{stlReader.cpp}
+\lstinputlisting[language=C++]{stlReader.cpp}
 \newpage

### 2018-03-31 stlReader.cpp

@@ -1,14 +1,13 @@
-// Name: Nicholas
-// Assignment: Final Project 
+// Name: Nicholas Czaban
+// Assignment: Final Project
 // Title: STL File Reader
 // Course: CS 200 
 // Semester: Spring year
 // Instructor: D. Byrnes
 // Date: 2/19/18
 // Program description: Reads STL files created by OpenSCAD and creates a vector of Facets
-// Known Bugs: The reader can only handle STL facets with three vertices and specific headings as provided by OpenSCAD. It is unclear whether or not every STL file generated by OpenSCAD or other programs will use the same format
-// Creativity:
-// Instructions: With this file in your working directory, use a #include "stlReader.cpp" in your program and call the functions below
+// Known Bugs: The reader can only handle STL facets with three vertices and specific headings as provided by OpenSCAD. It is unclear whether or not every STL file generated by other programs will use the same format
+// Instructions: Use a #include "stlReader.cpp" in your program and call the functions below
 
 #include <fstream>
 #include <iostream>
@@ -86,7 +85,8 @@ std::vector<Facet> readFile(std::string filename){
   return facets;
 }
 
-/* main() used to test functionality of stlReader, is not needed for anything else
+// main() used to test functionality of stlReader, is not needed for anything else
+/* 
 int main(){
   std::vector<Facet> facets = readFile("test.stl");
   for(int i = 0; i<facets.size(); i++){

### 2018-03-31 Paper.tex
Explained scad syntax

--- 2018-03-31 Paper.tex
@@
 \addtolength{\topmargin}{-0.5in}
+
+\lstset{language=C++, upquote=true, breaklines=true}
+
+\lstdefinelanguage{scad}
+{
+  basicstyle={\small\ttfamily},
+  morekeywords={true, false},
+  sensitive=true,
+  morecomment=[l][//],
+  commentstyle=\it,
+  morestring=[b]''
+  breaklines=true,
+  showstringspaces=false,
+}
 
@@
 \subsection{General Concepts}
-\indent Constructive Solid Geometry (CSG) is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
+Constructive Solid Geometry (CSG) is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
 The available boolean operations include unions, intersections, and differences; the types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations)\cite{cs_dictionary}.
+This technique provides a simpler way to create three-dimensional shapes without the need to define each vertex
 \subsubsection{Real-Time Rendering and Image-Based CSG}
-\indent One substantial category in CSG is the difference between Real-Time Rendering of CSG objects and Image-Based CSG objects.
+One substantial category in CSG is the difference between Real-Time Rendering of CSG objects and Image-Based CSG objects.
 Real-time rendering poses a number of issues, particularly for shapes that are interactively modified during runtime.
@@
 \subsection{CSG Algorithms}
-\indent Given the rendering complexity of constructive solid geometry, several techniques have been developed to improve the speed and quality of CSG primitives.
+Given the rendering complexity of constructive solid geometry, several techniques have been developed to improve the speed and quality of CSG primitives.
 In Nicholas Wilt's implementation of CSG in the Object Oriented Ray Tracing (OORT) class library, complicated calculations can be short-circuited by creating optimized CSG trees where simpler objects are the left children of each boolean operation.
@@
 \subsection{Display Program}
-\subsection{Asteroids}
 \subsubsection{CSG Usage}
@@
 The shapes are exported as .stl files, which can be read using the reader function located in \texttt{stlReader.cpp}.
+The code within a standard .scad file will look similar to this:
+\lstset{language=scad}
+\begin{lstlisting}
+  // A 6-sided die with rounded corners and circular faces
+  intersection() {
+    cube(15, center=true);
+    rotate(90, [1,0,0]) sphere(r=10, \$fn=50);
+  }
+\end{lstlisting}
+The \texttt{intersection()} function takes no inputs in its declaration, instead operating on the objects created within its code block.
+\texttt{union()} and \texttt{difference()} work in the same way.
+The \texttt{cube()} function can either create a regular cube when passed a single number, as above, or create rectangular prisms with user-defined dimensions.
+Instead of a single number, the user would pass the dimensions as such: \texttt{cube([15,10,12], center=true);}.
+The \texttt{center=true} parameter sets the cube with its center at the origin; a parameter of \texttt{center=false} would draw the cube with its bottom corner at the origin.
+Spheres\\
+
+\noindent
 \subsubsection{OpenGL Implementation}
+\subsection{Asteroids}
 \section{Conclusion}
-
+\newpage
+\section{Appendix}
+\subsection{stlReader.cpp}
+\lstinputlisting[language=C++]{stlReader.cpp}
 \newpage

### 2018-04-02 Paper.tex
Fixed gluLookAt and other minor edits

@@
 {
+  morecomment=[l][//],
+  commentstyle={\it},
   basicstyle={\small\ttfamily},
@@
   sensitive=true,
-  morecomment=[l][//],
-  commentstyle=\it,
   morestring=[b]''
@@
 \begin{lstlisting}
-  // A 6-sided die with rounded corners and circular faces
   intersection() {
@@
 The \texttt{center=true} parameter sets the cube with its center at the origin; a parameter of \texttt{center=false} would draw the cube with its bottom corner at the origin.
-Spheres\\
+Cylinders have independantly defined radii for both ends; a cone can be created by setting one radius to 0.
+As seen in the \texttt{rotate} call, vectors are contained within square brackets.For the sake of simplicity, and to better explore the possibilities of CSG, only rectangular prisms, cones, cylinders, and spheres were used to build each object.
+For transformations, only rotate and translate were used.\\
 
-\noindent
+\noindent In order to 
 \subsubsection{OpenGL Implementation}

### 2018-04-02 game.cpp

@@ -32,7 +32,7 @@ static GLfloat zpos = 0;
 static GLfloat speed = 0.2;
 
 GLfloat light_specular[] = {1, 1, 1, 1.0};
-GLfloat light_position[] = {0.0, -2.0, -10.0, 0.0};
+GLfloat light_position[] = {0.0, 3.0, -5.0, 0.0};
 
 std::vector<Facet> asteroids = readFile("pipes.stl");
 std::vector<Facet> spaceship = readFile("spaceship.stl");
@@ -51,12 +51,42 @@ void test_draw(){
   glEnd();
 }
 
+void flight(){
+  
+  // Change spaceship position
+  zpos += speed;
+  if(zpos <= 100){
+	zpos = 0;
+	speed += 0.01;
+  }
+  if(leftKey){
+	if(xpos > -10) xpos -= 0.2;
+	else xpos = -10;
+  }
+  if(rightKey){
+	if(xpos < 10) xpos += 0.2;
+	else xpos = 10;
+  }
+  if(upKey){
+	if(ypos < 10) ypos += 0.2;
+	else ypos = 10;	
+  }
+  if(downKey){
+	if(ypos > -10) ypos -= 0.2;
+	else ypos = -10;
+  }
+  glutPostRedisplay();
+}
+
+
 void setup_scene(){
   // Draw the spaceship
-  glPushMatrix();
-  glRotatef(-1*angle2, 1.0, 0.0, 0.0);
-  glRotatef(-1*angle, 0.0, 1.0, 0.0);
 
+  gluLookAt(0,-1,5,
+			0,0,0,
+			0,1,0);
+
+  glPushMatrix();
   glColor3f(0.4, 0.4, 0.4);
   glTranslatef(xpos, ypos, 1);
   if(leftKey) glRotatef(15, 0, 0, 1);
@@ -71,26 +101,12 @@ void setup_scene(){
   
   glBegin(GL_TRIANGLES);
   for(int i = 0; i<spaceship.size(); i++){
-	glNormal3f(spaceship[i].normal[0], -1*spaceship[i].normal[1], spaceship[i].normal[2]);
+	glNormal3f(spaceship[i].normal[0], spaceship[i].normal[1], spaceship[i].normal[2]);
 	if((spaceship[i].normal[2] == -1) && (spaceship[i].vertices[0][2] < -1.5)) glColor3f(.4, .6, 1);
[…]

### 2018-04-03 Paper.tex
Explained stlReader.cpp

@@
 Cylinders have independantly defined radii for both ends; a cone can be created by setting one radius to 0.
-As seen in the \texttt{rotate} call, vectors are contained within square brackets.For the sake of simplicity, and to better explore the possibilities of CSG, only rectangular prisms, cones, cylinders, and spheres were used to build each object.
+As seen in the \texttt{rotate} call, vectors are contained within square brackets.
+For the sake of simplicity, and to better explore the possibilities of CSG, only rectangular prisms, cones, cylinders, and spheres were used to build each object.
 For transformations, only rotate and translate were used.\\
 
-\noindent In order to 
+\noindent In order to manipulate the objects created with OpenSCAD, the CSG objects were exported in .stl files. Each file is formatted in the same way:
+\begin{lstlisting}
+  facet normal 0.603056 0.409837 0.684366
+    outer loop
+      vertex 3.33769 2.42497 3.63721
+      vertex 3.14634 2.28595 3.88909
+      vertex 3.48337 2.21061 3.63721
+    endloop
+  endfacet
+\end{lstlisting}
+The object is broken down into a series of triangles and a normal vector is calculated.
+Using this pattern, the .stl files were read by the helper program, stlReader.cpp, and fed into OpenGL's \texttt{GL{\bf\_}TRIANGLES} primitive to rebuild the object in a more interactive environment.\\
+
+\noindent stlReader.cpp uses the struct \texttt{Facet} to store the data for each triangle, with each \texttt{Facet} being added to a vector containing all the pieces of the shape.
+The helper program is currently only configured to read .stl files that use the ASCII-style layout, and cannot read binary .stl files.\\
+
+\noindent In total, seven objects were created using OpenSCAD. These include, but are not limited to, a die, a taurus, an asteroid, a spaceship, and an abstract cookie-cutter.
 \subsubsection{OpenGL Implementation}

### 2018-04-03 game.cpp

@@ -34,28 +34,15 @@ static GLfloat speed = 0.2;
 GLfloat light_specular[] = {1, 1, 1, 1.0};
 GLfloat light_position[] = {0.0, 3.0, -5.0, 0.0};
 
-std::vector<Facet> asteroids = readFile("pipes.stl");
+std::vector<Facet> asteroids = readFile("asteroid.stl");
 std::vector<Facet> spaceship = readFile("spaceship.stl");
 
 bool upKey, leftKey, rightKey, downKey;
 
-void test_draw(){
-  gluLookAt(0,0,-5,
-			0,0,0,
-			0,1,0);
-  glColor3f(1,1,1);
-  glBegin(GL_POLYGON);
-  glVertex3f(-1,-1,0);
-  glVertex3f(0,1,0);
-  glVertex3f(1,-1,0);
-  glEnd();
-}
-
 void flight(){
-  
   // Change spaceship position
   zpos += speed;
-  if(zpos <= 100){
+  if(zpos >= 100){
 	zpos = 0;
 	speed += 0.01;
   }
@@ -78,14 +65,13 @@ void flight(){
   glutPostRedisplay();
 }
 
-
 void setup_scene(){
-  // Draw the spaceship
-
   gluLookAt(0,-1,5,
 			0,0,0,
 			0,1,0);
 
+  glRotatef(-25,0,1,0);
+  // Draw the spaceship
   glPushMatrix();
   glColor3f(0.4, 0.4, 0.4);
   glTranslatef(xpos, ypos, 1);
@@ -93,11 +79,6 @@ void setup_scene(){
   if(rightKey) glRotatef(-15, 0, 0, 1);
   if(upKey) glRotatef(-15, 1, 0, 0);
   if(downKey) glRotatef(15, 1, 0, 0);
-
-  /****
-	   The stl files generated by OpenSCAD use the right-hand rule for its coordinate system.
-	   Thus, in order to correctly render the shapes, the z-coordinates are multiplied by -1.
-   ****/
   
   glBegin(GL_TRIANGLES);
   for(int i = 0; i<spaceship.size(); i++){
@@ -107,6 +88,20 @@ void setup_scene(){
 	  glVertex3f(spaceship[i].vertices[j][0], spaceship[i].vertices[j][1], spaceship[i].vertices[j][2]);
 	}
[…]

### 2018-04-03 stlReader.cpp

@@ -6,7 +6,7 @@
 // Instructor: D. Byrnes
 // Date: 2/19/18
 // Program description: Reads STL files created by OpenSCAD and creates a vector of Facets
-// Known Bugs: The reader can only handle STL facets with three vertices and specific headings as provided by OpenSCAD. It is unclear whether or not every STL file generated by other programs will use the same format
+// Known Bugs: The reader can only handle ASCII-style STL files. It cannot read binary-style STL files.
 // Instructions: Use a #include "stlReader.cpp" in your program and call the functions below
 
 #include <fstream>

### 2018-04-03 Paper.tex
Started CSG Algorithm section and fixed spacing

@@
 \usepackage{listings}
-\linespread{1.25}
 \usepackage[letterpaper, margin=1in]{geometry}
@@
 \maketitle
-
+\begin{doublespace}
 \section{Constructive Solid Geometry}
@@
 \subsection{CSG Algorithms}
-Given the rendering complexity of constructive solid geometry, several techniques have been developed to improve the speed and quality of CSG primitives.
+Given the rendering complexity of constructive solid geometry, several algorithms have been developed to improve the speed and quality of CSG primitives.
+These include the Goldfeather Algorithm, the layered Goldfeather algorithm, and Sequenced Convex Subtraction (SCS) algorithm\cite{hardware_csg}.
+Goldfeather's algorithm quickly became the precursor to all image-based CSG algorithms.
+\\
+
+There are other, more minute techniques that can improve performance when rendering CSG objects.
 In Nicholas Wilt's implementation of CSG in the Object Oriented Ray Tracing (OORT) class library, complicated calculations can be short-circuited by creating optimized CSG trees where simpler objects are the left children of each boolean operation.
 In OORT, ray tracing calculations between two objects on a CSG tree always test against the left object first.
-Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE\cite{raytrace}.
+Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE\cite{raytrace}.\\
 \section{Software}
 \subsection{Display Program}
+The first implementation of constructive solid geometry is a simple display program.
+It takes as a command-line argument an .stl file, containing the vertices of the shape to be rendered in OpenGL.
+The shape can be rotated on its axis by clicking and dragging the shape.
 \subsubsection{CSG Usage}
@@
 \newpage
+\end{doublespace}
+\begin{onehalfspace}
 \section{Appendix}
@@
 \lstinputlisting[language=C++]{stlReader.cpp}
+\subsection{display.cpp}
+\lstinputlisting[language=C++]{display.cpp}
 \newpage
@@
 \end{thebibliography}
+\end{onehalfspace}
 \end{document}

### 2018-04-03 display.cpp

@@ -18,7 +18,6 @@
 // Sources consulted: any books consulted in writing the program
 // Program description: Using the function in stlReader.cpp, create a model of a CSG object with OpenGL's TRIANGLES primitive.
 // Known bugs: Program does not check to see if the file is of type STL before running. On occasion, some triangles on the shape will not be drawn correctly; this issue is resolved on repeat executions.
-// Creativity: 
 // Instructions: Compile and run on the command line. Type the name of the STL file as the first command line option when running. Click and move the mouse to rotate the object in space. Press 'q' to exit the program.
 
 GLint window_width = 500;

### 2018-04-04 Paper.tex
Add information about Goldfeather algorithm

@@
 The available boolean operations include unions, intersections, and differences; the types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations)\cite{cs_dictionary}.
-This technique provides a simpler way to create three-dimensional shapes without the need to define each vertex
+This technique provides a simpler way to create three-dimensional shapes without the need to define each individual vertex or face.\\
+
+Constructive Solid Geometry are organized in binary trees, where leaf nodes are the solid primitives and interior nodes are the boolean operations.
+The object is then created recursively, with each boolean operation being replaced by the object which it created\cite{gold}.
 \subsubsection{Real-Time Rendering and Image-Based CSG}
@@
 Given the rendering complexity of constructive solid geometry, several algorithms have been developed to improve the speed and quality of CSG primitives.
-These include the Goldfeather Algorithm, the layered Goldfeather algorithm, and Sequenced Convex Subtraction (SCS) algorithm\cite{hardware_csg}.
-Goldfeather's algorithm quickly became the precursor to all image-based CSG algorithms.
+These include the Goldfeather Algorithm, the layered Goldfeather algorithm, and Sequenced Convex Subtraction (SCS) algorithm.
+The Goldfeather algorithm quickly became the precursor to all image-based CSG algorithms\cite{hardware_csg}.\\
+
+Goldfeather's algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effectively use the z-buffer on supported graphics hardware\cite{hardware_csg}.
+The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm, which traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers\cite{gold}.
+Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates.
 \\
 
-There are other, more minute techniques that can improve performance when rendering CSG objects.
+In their extension of Goldfeather's algorithm, Stewart et.
+al.
+exploit instances where the maximum depth complexity $k$ (the number of primitives covering each pixel) is less than the number of primitives $n$ in a normalized CSG tree\cite{layered_gold}.\\
+
+In addition to the core algorithms, there are other, more minute techniques that can improve performance when rendering CSG objects.
 In Nicholas Wilt's implementation of CSG in the Object Oriented Ray Tracing (OORT) class library, complicated calculations can be short-circuited by creating optimized CSG trees where simpler objects are the left children of each boolean operation.
 In OORT, ray tracing calculations between two objects on a CSG tree always test against the left object first.
-Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE\cite{raytrace}.\\
+Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE\cite{raytrace}.
 \section{Software}
@@
 
-\noindent In order to manipulate the objects created with OpenSCAD, the CSG objects were exported in .stl files. Each file is formatted in the same way:
+In order to manipulate the objects created with OpenSCAD, the CSG objects were exported in .stl files. Each file is formatted in the same way:
 \begin{lstlisting}
@@
 The object is broken down into a series of triangles and a normal vector is calculated.
-Using this pattern, the .stl files were read by the helper program, stlReader.cpp, and fed into OpenGL's \texttt{GL{\bf\_}TRIANGLES} primitive to rebuild the object in a more interactive environment.\\
-
-\noindent stlReader.cpp uses the struct \texttt{Facet} to store the data for each triangle, with each \texttt{Facet} being added to a vector containing all the pieces of the shape.
+Using this pattern, the .stl files were read by the helper program, stlReader.cpp.
+stlReader.cpp uses the struct \texttt{Facet} to store the data for each triangle, with each \texttt{Facet} being added to a vector containing all the pieces of the shape.
 The helper program is currently only configured to read .stl files that use the ASCII-style layout, and cannot read binary .stl files.\\
 
-\noindent In total, seven objects were created using OpenSCAD. These include, but are not limited to, a die, a taurus, an asteroid, a spaceship, and an abstract cookie-cutter.
+In total, seven objects were created using OpenSCAD. These include, but are not limited to, a die, a taurus, an asteroid, a spaceship, and an abstract cookie-cutter.
 \subsubsection{OpenGL Implementation}
+Once the .stl file is parsed, the vector of \texttt{Facet}s is fed into OpenGL's \texttt{GL{\bf\_}TRIANGLES} primitive to rebuild the object in a more interactive environment. The display program 
 \subsection{Asteroids}
@@
 \bibitem{cs_dictionary}
-  Andrew Butterfield and Gerard Ngondi,
+  Butterfield, Andrew and Gerard Ngondi,
   \textit{A Dictionary of Computer Science},
@@
   2016
+
+\bibitem{gold}
+  Goldfeather, Jack, Jeff Hultquist, and Henry Fuchs. ``Fast Constructive Solid Geometry in the Pixel-Powers Graphics System.'' {\it Computer Graphics, SIGGRAPH '86 Proceedings}, vol. 20, no. 4, 1986,
+pp. 107-116
 
@@
   Ooms, Kristien. ``A 3D inclusion test on large dataset.'' {\it Developments in 3D Geo-Information Sciences}. Edited by Philippe De Maeyer and Tijs Neutens. Springer, New York, 2010.
+
+\bibitem{layered_gold}
+  Stewart, Nigel, Geoff Leach, and Sabu John. ``An Improved Z-Buffer CSG Rendering Algorithm.'' {\it 1998 Eurographics / SIGGRAPH Workshop on Graphics Hardware}, ACM, 25-30, 1998.
 

### 2018-04-04 display.cpp

@@ -18,7 +18,7 @@
 // Sources consulted: any books consulted in writing the program
 // Program description: Using the function in stlReader.cpp, create a model of a CSG object with OpenGL's TRIANGLES primitive.
 // Known bugs: Program does not check to see if the file is of type STL before running. On occasion, some triangles on the shape will not be drawn correctly; this issue is resolved on repeat executions.
-// Instructions: Compile and run on the command line. Type the name of the STL file as the first command line option when running. Click and move the mouse to rotate the object in space. Press 'q' to exit the program.
+// Instructions: Compile and run on the command line. Type the name of the STL file as the first command line argument when running. Click and move the mouse to rotate the object in space. Press 'q' to exit the program.
 
 GLint window_width = 500;
 GLint window_height = 500;

### 2018-04-04 Paper.tex
Explained OpenGL in paper

@@
 
-In total, seven objects were created using OpenSCAD. These include, but are not limited to, a die, a taurus, an asteroid, a spaceship, and an abstract cookie-cutter.
+In total, seven objects were created using OpenSCAD. These include, but are not limited to: a die, a torus, an asteroid, a spaceship, and an abstract cookie-cutter.
 \subsubsection{OpenGL Implementation}
-Once the .stl file is parsed, the vector of \texttt{Facet}s is fed into OpenGL's \texttt{GL{\bf\_}TRIANGLES} primitive to rebuild the object in a more interactive environment. The display program 
+Once the .stl file is parsed, the vector of \texttt{Facet}s is fed into OpenGL's \texttt{GL{\bf\_}TRIANGLES} primitive to rebuild the object in a more interactive environment.
+The display program uses a white specular point light and a material shininess of 50.
+A \texttt{mouse} and \texttt{motion} function enable the user to rotate the object on its axis, with the rotations being performed within the \texttt{display} function.
+Within the \texttt{glBegin} and \texttt{glEnd} block, the vector \texttt{newShape} is scanned, with each \texttt{Facet}'s normal vector and vertices passed to the respective OpenGL command.
+In order to correctly light the objects, the vertices are passed in the reverse order they were printed in the .stl file.
 \subsection{Asteroids}
@@
 \section{Appendix}
-\subsection{stlReader.cpp}
-\lstinputlisting[language=C++]{stlReader.cpp}
-\subsection{display.cpp}
-\lstinputlisting[language=C++]{display.cpp}
+\subsection{OpenSCAD}
+\subsubsection{die.scad}
+\subsubsection{connector.scad}
+\subsubsection{torus.scad}
+\subsubsection{asteroid.scad}
+\subsection{C++}
+\subsubsection{stlReader.cpp}
+%\lstinputlisting[language=C++]{stlReader.cpp}
+\subsubsection{display.cpp}
+%\lstinputlisting[language=C++]{display.cpp}
 \newpage

### 2018-04-04 display.cpp

@@ -43,7 +43,6 @@ void display(){
   glPushMatrix();
   glRotatef(-1*angle2, 1.0, 0.0, 0.0);
   glRotatef(-1*angle, 0.0, 1.0, 0.0);
-  glColor3f(1.0, 0.0, 0.0);
   glBegin(GL_TRIANGLES);
   for(int i = 0; i<newShape.size(); i++){
 	glNormal3f(newShape[i].normal[0], newShape[i].normal[1], newShape[i].normal[2]);

### 2018-04-04 Paper.tex
Up to 5 pages on the paper

@@
 
-\lstset{language=C++, upquote=true, breaklines=true}
+\lstset{language=C++, upquote=true, breaklines=true, showstringspaces=false}
 
@@
 The \texttt{center=true} parameter sets the cube with its center at the origin; a parameter of \texttt{center=false} would draw the cube with its bottom corner at the origin.
-Cylinders have independantly defined radii for both ends; a cone can be created by setting one radius to 0.
+Cylinders have independently defined radii for both ends; a cone can be created by setting one radius to 0.
 As seen in the \texttt{rotate} call, vectors are contained within square brackets.
@@
 
-In total, seven objects were created using OpenSCAD. These include, but are not limited to: a die, a torus, an asteroid, a spaceship, and an abstract cookie-cutter.
+In total, seven objects were created using OpenSCAD. These include, but are not limited to: a die, a torus, an asteroid, a spaceship, and a 6-sided cookie-cutter.
 \subsubsection{OpenGL Implementation}
@@
 The display program uses a white specular point light and a material shininess of 50.
-A \texttt{mouse} and \texttt{motion} function enable the user to rotate the object on its axis, with the rotations being performed within the \texttt{display} function.
+A \texttt{mouse} and \texttt{motion} function enable the user to rotate the object on its axis, with the rotations being performed within the \texttt{display} function; the user can also quit the program with the 'q' key.
 Within the \texttt{glBegin} and \texttt{glEnd} block, the vector \texttt{newShape} is scanned, with each \texttt{Facet}'s normal vector and vertices passed to the respective OpenGL command.
-In order to correctly light the objects, the vertices are passed in the reverse order they were printed in the .stl file.
+OpenSCAD uses a right-handed coordinate system while OpenGL uses a left-handed system; thus, the vertices are passed in the reverse order they were printed in the .stl file in order to correctly light the objects.
 \subsection{Asteroids}
+The second implementation of constructive solid geometry is an interactive game, where the user controls a spaceship and avoids asteroids.
+In comparison to the display program, this program enables the \texttt{GL{\bf\_}COLOR{\bf\_}MATERIAL} state in its \texttt{init} call, which allows the \texttt{glColor} function to be used in conjunction with lighting.
+The spaceship surface is colored a light grey, with the engine trails colored blue; the asteroids are colored red.
+The game removes the mouse-controlled camera in place of a static camera set with the \texttt{gluLookAt} function.
+Additionally, more key bindings are added to control the movement of the spaceship.
+Each of the WASD keys has a respective boolean variable that is set to true when the key is pressed and false when the key is released.
+This method allows for diagonal movement as well as cardinal movement.
 \section{Conclusion}
+
 \newpage
@@
 \begin{onehalfspace}
-\section{Appendix}
+\section{Source Code}
 \subsection{OpenSCAD}
@@
 \subsubsection{asteroid.scad}
+\subsubsection{pipes.scad}
+\subsubsection{spaceship.scad}
+\subsubsection{cookie\_cutter.scad}
+
 \subsection{C++}
 \subsubsection{stlReader.cpp}
-%\lstinputlisting[language=C++]{stlReader.cpp}
+% \lstinputlisting[language=C++]{stlReader.cpp}
 \subsubsection{display.cpp}
-%\lstinputlisting[language=C++]{display.cpp}
+% \lstinputlisting[language=C++]{display.cpp}
+\subsubsection{game.cpp}
+% \lstinputlisting[language=C++]{game.cpp}
+
 \newpage

### 2018-04-04 display.cpp

@@ -15,7 +15,6 @@
 // Semester: Spring year
 // Instructor: D. Byrnes
 // Date: 2/19/18
-// Sources consulted: any books consulted in writing the program
 // Program description: Using the function in stlReader.cpp, create a model of a CSG object with OpenGL's TRIANGLES primitive.
 // Known bugs: Program does not check to see if the file is of type STL before running. On occasion, some triangles on the shape will not be drawn correctly; this issue is resolved on repeat executions.
 // Instructions: Compile and run on the command line. Type the name of the STL file as the first command line argument when running. Click and move the mouse to rotate the object in space. Press 'q' to exit the program.

### 2018-04-04 game.cpp

@@ -66,9 +66,7 @@ void flight(){
 }
 
 void setup_scene(){
-  gluLookAt(0,-1,5,
-			0,0,0,
-			0,1,0);
+  gluLookAt(0, -1, 5,    0, 0, 0,     0,1,0);
 
   glRotatef(-25,0,1,0);
   // Draw the spaceship

### 2018-04-06 Paper.tex
Explain Goldfeather algorithm

@@
 Given the rendering complexity of constructive solid geometry, several algorithms have been developed to improve the speed and quality of CSG primitives.
-These include the Goldfeather Algorithm, the layered Goldfeather algorithm, and Sequenced Convex Subtraction (SCS) algorithm.
-The Goldfeather algorithm quickly became the precursor to all image-based CSG algorithms\cite{hardware_csg}.\\
+These include the Goldfeather Algorithm, the layered Goldfeather algorithm, and Sequenced Convex Subtraction (SCS) algorithm.\\
 
-Goldfeather's algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effectively use the z-buffer on supported graphics hardware\cite{hardware_csg}.
+Goldfeather's algorithm quickly became the precursor to all image-based CSG algorithms\cite{hardware_csg}.
+The algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effectively use the z-buffer on supported graphics hardware\cite{hardware_csg}.
 The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm, which traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers\cite{gold}.
 Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates.
-\\
+When perfoming CSG operations in this algorithm, the visible section of the first object is stored in a z-buffer, denoted in Goldfeather's paper as \texttt{ZMIN}, and the visible section of the second object is stored in another z-buffer, denoted \texttt{ZTEMP}.
+Then, for each pixel in the screen, if the value of \texttt{ZTEMP} is less than (i.e.
+closer to the screen) \texttt{ZMIN}, a flag \texttt{F1} is set to true.\\
 
@@
 The spaceship surface is colored a light grey, with the engine trails colored blue; the asteroids are colored red.
+For the engine color, the code checks to see if the normal vector is pointing in the same direction as the engines, as well as if the z-value of the triangle's first vertex is at -1.5, at the back of the ship.
 The game removes the mouse-controlled camera in place of a static camera set with the \texttt{gluLookAt} function.

### 2018-04-07 Paper.tex
Fix perspective, explain 3 cases of Goldfeather, other minor paper edits

@@
 \usepackage{courier}
+\usepackage{graphicx}
+\graphicspath{ {images/} }
 \usepackage{setspace}
@@
 The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm, which traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers\cite{gold}.
-Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates.
-When perfoming CSG operations in this algorithm, the visible section of the first object is stored in a z-buffer, denoted in Goldfeather's paper as \texttt{ZMIN}, and the visible section of the second object is stored in another z-buffer, denoted \texttt{ZTEMP}.
-Then, for each pixel in the screen, if the value of \texttt{ZTEMP} is less than (i.e.
-closer to the screen) \texttt{ZMIN}, a flag \texttt{F1} is set to true.\\
+Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates.\\
+
+When perfoming CSG unions in this algorithm, the visible section of the first object is stored in a z-buffer, denoted in Goldfeather's paper as \texttt{ZMIN}, and the visible section of the second object is stored in another z-buffer, denoted \texttt{ZTEMP}.
+For all the pixels in the visible area of the second object, a flag \texttt{F1} is set to true.
+Then, for any pixels where \texttt{F1} is set, if \texttt{ZTEMP} greater than \texttt{ZMIN} (i.e.
+the second object is deeper in model space than the first object), the \texttt{F1} flag is cleared.
+The remaining pixels with \texttt{F1} set represent areas where the second object is visible from behind the first object.
+Both objects are then drawn without any overlap.\\
+
+In difference operations, the flag \texttt{F1} is set for all pixels of the first object and the front faces are stored in \texttt{ZTEMP}.
+If the \texttt{ZTEMP} values lie between the front and back of the second object, \texttt{F1} is cleared.
+The remaining \texttt{F1} pixels are not within the second object, but are on the front of the first object; the corresponding values of \texttt{ZTEMP} are passed into \texttt{ZMIN}.\\
+
+For intersection problems, \texttt{F1} is set for the front of the first object, then cleared if the value of \texttt{ZTEMP} is not between the front and back of the second object.
+Then, the same comparison is made with the second object stored in \texttt{ZTEMP} and compared between the front and back faces of the first object.
+Only the pixels with the flag set for both sets are drawn.\\
 
@@
 Within the \texttt{glBegin} and \texttt{glEnd} block, the vector \texttt{newShape} is scanned, with each \texttt{Facet}'s normal vector and vertices passed to the respective OpenGL command.
-OpenSCAD uses a right-handed coordinate system while OpenGL uses a left-handed system; thus, the vertices are passed in the reverse order they were printed in the .stl file in order to correctly light the objects.
+In order to correctly light the triangle elements, the vertices are passed in the reverse order of how were printed in the .stl file.
 \subsection{Asteroids}
@@
 For the engine color, the code checks to see if the normal vector is pointing in the same direction as the engines, as well as if the z-value of the triangle's first vertex is at -1.5, at the back of the ship.
-The game removes the mouse-controlled camera in place of a static camera set with the \texttt{gluLookAt} function.
+The game removes the mouse-controlled camera in place of a camera set with the \texttt{gluLookAt} function to constantly be directly behind the spaceship.
 Additionally, more key bindings are added to control the movement of the spaceship.
@@
 This method allows for diagonal movement as well as cardinal movement.
+Although the display program used \texttt{glOrtho} to set up the projection, the game uses \texttt{gluProjection}.
+This change was needed in order to simulate the correct depth perception on the obstacles.\\
+
+In the idle loop of the game, the position values of the spaceship are changed.
+If one of the directional booleans is true, \texttt{xpos} or \texttt{ypos} is changed accordingly.
+When the scene is redrawn, the camera position uses \texttt{xpos} and \texttt{ypos} to remain behind the ship.
+Another variable, \texttt{zpos} is also changed; this variable does not tranlate the spaceship itself, but instead translates all other obstacles onscreen.
+The \texttt{zpos} variable is cyclical; once the value is less than -100, \texttt{zpos} is reset to 0, and the \texttt{speed}, \texttt{zpos}'s rate of change, is increased.\\
+
+The increased movement of the game objects highlighted one major challenge of converting the OpenSCAD designs to OpenGL.
+OpenSCAD uses a right-handed coordinate system, while OpenGL uses a left-handed system.
+As the objects moved towards and away from the camera, parts of the objects that should have been hidden were visible, and vice versa.
+For example, the spaceship engines jut out from the back of the ship.
+Thus, a viewer positioned in between the two engines should be able to see the right curve of the left engine and the left curve of the right engine, as shown below left.
+However, the viewer instead could see the left curve of the left engine and the right curve of the right engine, as seen below right; this perspective made the engines appear recessed instead of external.
+In order to solve this issue, the z-coordinate of each vertex was multiplied by -1.
+\begin{figure}[h]
+  \includegraphics[width=8cm]{GoodPersp}
+  \includegraphics[width=8cm]{BadPersp}
+  \centering
+\end{figure}
+
+With the exception of the spaceship, the other objects are given random numbers to determine their position on the screen and their rotational speed. 
 \section{Conclusion}
@@
 \begin{onehalfspace}
-\section{Source Code}
-\subsection{OpenSCAD}
-\subsubsection{die.scad}
-\subsubsection{connector.scad}
-\subsubsection{torus.scad}
-\subsubsection{asteroid.scad}
-\subsubsection{pipes.scad}
-\subsubsection{spaceship.scad}
-\subsubsection{cookie\_cutter.scad}
+\section*{Source Code}
+\subsection*{OpenSCAD}
+\subsubsection*{die.scad}
+\subsubsection*{connector.scad}
+\subsubsection*{torus.scad}
+\subsubsection*{asteroid.scad}
+\subsubsection*{pipes.scad}
+\subsubsection*{spaceship.scad}
+\subsubsection*{cookie\_cutter.scad}
 
-\subsection{C++}
-\subsubsection{stlReader.cpp}
+\subsection*{C++}
+\subsubsection*{stlReader.cpp}
 % \lstinputlisting[language=C++]{stlReader.cpp}
-\subsubsection{display.cpp}
+\subsubsection*{display.cpp}
 % \lstinputlisting[language=C++]{display.cpp}
-\subsubsection{game.cpp}
+\subsubsection*{game.cpp}
 % \lstinputlisting[language=C++]{game.cpp}

### 2018-04-07 game.cpp

@@ -35,14 +35,15 @@ GLfloat light_specular[] = {1, 1, 1, 1.0};
 GLfloat light_position[] = {0.0, 3.0, -5.0, 0.0};
 
 std::vector<Facet> asteroids = readFile("asteroid.stl");
+std::vector<Facet> pipe = readFile("pipes.stl");
 std::vector<Facet> spaceship = readFile("spaceship.stl");
 
 bool upKey, leftKey, rightKey, downKey;
 
 void flight(){
   // Change spaceship position
-  zpos += speed;
-  if(zpos >= 100){
+  zpos -= speed;
+  if(zpos <= -100){
 	zpos = 0;
 	speed += 0.01;
   }
@@ -66,39 +67,55 @@ void flight(){
 }
 
 void setup_scene(){
-  gluLookAt(0, -1, 5,    0, 0, 0,     0,1,0);
+  glTranslatef(0,0,-10);
+  gluLookAt(xpos, ypos+1, 5,    xpos, ypos, 0,     0,1,0);
 
-  glRotatef(-25,0,1,0);
   // Draw the spaceship
   glPushMatrix();
   glColor3f(0.4, 0.4, 0.4);
   glTranslatef(xpos, ypos, 1);
   if(leftKey) glRotatef(15, 0, 0, 1);
   if(rightKey) glRotatef(-15, 0, 0, 1);
-  if(upKey) glRotatef(-15, 1, 0, 0);
-  if(downKey) glRotatef(15, 1, 0, 0);
+  if(upKey) glRotatef(15, 1, 0, 0);
+  if(downKey) glRotatef(-15, 1, 0, 0);
   
   glBegin(GL_TRIANGLES);
   for(int i = 0; i<spaceship.size(); i++){
 	glNormal3f(spaceship[i].normal[0], spaceship[i].normal[1], spaceship[i].normal[2]);
 	if((spaceship[i].normal[2] == -1) && (spaceship[i].vertices[0][2] < -1.5)) glColor3f(.4, .6, 1);
 	for(int j = 2; j>=0; j--){
-	  glVertex3f(spaceship[i].vertices[j][0], spaceship[i].vertices[j][1], spaceship[i].vertices[j][2]);
+	  glVertex3f(spaceship[i].vertices[j][0], spaceship[i].vertices[j][1], -1*spaceship[i].vertices[j][2]);
 	}
[…]

### 2018-04-07 Paper.tex
Wrote large segments of the paper

@@
 Given the rendering complexity of constructive solid geometry, several algorithms have been developed to improve the speed and quality of CSG primitives.
-These include the Goldfeather Algorithm, the layered Goldfeather algorithm, and Sequenced Convex Subtraction (SCS) algorithm.\\
+Two such algorithms are the Goldfeather Algorithm and the layered Goldfeather algorithm.
 
+\subsubsection{Goldfeather Algorithm}
 Goldfeather's algorithm quickly became the precursor to all image-based CSG algorithms\cite{hardware_csg}.
-The algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effectively use the z-buffer on supported graphics hardware\cite{hardware_csg}.
+The algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effectively use the z-buffer on supported graphics hardware.
 The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm, which traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers\cite{gold}.
@@
 Then, the same comparison is made with the second object stored in \texttt{ZTEMP} and compared between the front and back faces of the first object.
-Only the pixels with the flag set for both sets are drawn.\\
+Only the pixels with the flag set for both sets are drawn.
 
+\subsubsection{Layered Goldfeather Algorithm}
 In their extension of Goldfeather's algorithm, Stewart et.
 al.
-exploit instances where the maximum depth complexity $k$ (the number of primitives covering each pixel) is less than the number of primitives $n$ in a normalized CSG tree\cite{layered_gold}.\\
+exploit instances where the maximum depth complexity $k$ (the number of primitives covering each pixel) is less than the number of primitives $n$ in a normalized CSG tree\cite{layered_gold}.
+For example, a scene with five spheres in a line will have no overlapping objects when the viewer is orthogonal to the line.
+In this case, $k=1$ and is less than the number of primitives, $n=5$.
+If the viewer looked directly down the line, the depth complexity would equal the number of primitives and would not be suitable for the layered algorithm.\\
 
+The layered algorithm defines layers by their surface parity, or the number of edges in the z-buffer.
+Parity is stored as a flag in the algorithm; parity is even when outside an object and odd when inside an object.
+While generally effective, this definition places important constraints on the usable surfaces: the surface cannot self-intersect, and all volumes must be completely covered in a surface with no holes\cite{layered_gold}.\\
+
+The original Goldfeather algorithm performs a single clipping operation in $an+b$ time, where $a$ and $b$ are constants determined by polygon rasterization speed and z-buffer copying, respectively\cite{layered_gold}.
+Each primitive is clipped once, giving an upper runtime bound of $O(n^2)$.
+In the layered Goldfeather algorithm, each primitive is rendered once into the layer and once to clip the layer.
+Each layer is copied into the z-buffer, so the time to clip a single layer in the layered approach is $2an+b$.
+In total, with each layer clipped once, the layered algorithm runs in $2akn+bk$, or $O(kn)$\cite{layered_gold}.
+As stated above, the algorithm performs better than the original Goldfeather method when $k<n$.
+
+\subsubsection{Other Techniques}
 In addition to the core algorithms, there are other, more minute techniques that can improve performance when rendering CSG objects.
@@
 In OORT, ray tracing calculations between two objects on a CSG tree always test against the left object first.
-Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE\cite{raytrace}.
+Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE\cite{raytrace}.\\
+
+The Goldfeather algorithm uses two separate z-buffers, which can create issues on systems that do not support multiple z-buffers.
+Alternative approaches, such as copying one set of z-values into main memory, are inadequately slow and values can lose exactness during copying.
+One workaround is to save the second z-buffer into another, unused buffer, such as the color buffer.
+Commonly referred to as render-to-texture, color data is discarded and the texels of the new p-buffer have a one-to-one match to the pixels in the frame buffer\cite{hardware_csg}.\\
+
+In the layered Goldfeather approach, the biggest detriment to its speed is reading from the stencil buffer to calculate the maximum depth complexity.
+However, with modern graphics hardware the depth complexity can be calculated without reading directly from the stencil buffer by using extensions built in to NVidia and ATI hardware, such as NV\_occlusion\_query\cite{hardware_csg}.
+
 \section{Software}
@@
 In order to correctly light the triangle elements, the vertices are passed in the reverse order of how were printed in the .stl file.
-\subsection{Asteroids}
+\subsection{Asteroid Game}
 The second implementation of constructive solid geometry is an interactive game, where the user controls a spaceship and avoids asteroids.
@@
   \centering
+  \caption{Two different views of the spaceship model. The left model is the intended view, the right model is how the object was initially rendered in OpenGL.}
 \end{figure}
 
-With the exception of the spaceship, the other objects are given random numbers to determine their position on the screen and their rotational speed. 
+Three random floating-point variables between 0 and 1 are created on each cycle; these three variables determine the rotation and position of each obstacle onscreen.
+For the \texttt{glTranslate} calls, the random numbers are multiplied by 20 then subtracted by 10 to place it within the playing area; the \texttt{glRotate} calls use the random values as they are.
+Each of the three variables is shuffled in their usage in the objects, so that no object has the same rotation or position as another.
 \section{Conclusion}

### 2018-04-07 game.cpp

@@ -23,9 +23,6 @@
 
 GLint window_width = 700;
 GLint window_height = 700;
-static int moving = 0, startx = 0, starty = 0;
-static GLfloat angle = 0;
-static GLfloat angle2 = 180;
 static GLfloat xpos = 0;
 static GLfloat ypos = 0;
 static GLfloat zpos = 0;
@@ -37,13 +34,18 @@ GLfloat light_position[] = {0.0, 3.0, -5.0, 0.0};
 std::vector<Facet> asteroids = readFile("asteroid.stl");
 std::vector<Facet> pipe = readFile("pipes.stl");
 std::vector<Facet> spaceship = readFile("spaceship.stl");
+std::vector<Facet> connector = readFile("connector.stl");
 
 bool upKey, leftKey, rightKey, downKey;
+float r1, r2, r3;
 
 void flight(){
   // Change spaceship position
   zpos -= speed;
   if(zpos <= -100){
+	r1 = float(rand())/RAND_MAX;
+	r2 = float(rand())/RAND_MAX;
+	r3 = float(rand())/RAND_MAX;
 	zpos = 0;
 	speed += 0.01;
   }
@@ -93,12 +95,12 @@ void setup_scene(){
 
   // Draw a pipe
   glPushMatrix();
-  glTranslatef(0,0,-25-zpos);
-  glRotatef(-1.5*zpos, 0, 0, 0.8);
+  glTranslatef((20*r3)-10, (20*r1)-10, -25-zpos);
+  glRotatef(-1.5*zpos, r1, r2, r3);
   glColor3f(0.6, 0.6, 0.6);
   glBegin(GL_TRIANGLES);
   for(int i = 0; i<pipe.size(); i++){
-	glNormal3f(pipe[i].normal[0], pipe[i].normal[1], pipe[i].normal[2]);
+	glNormal3f(-1*pipe[i].normal[0], pipe[i].normal[1], pipe[i].normal[2]);
 	for(int j = 2; j>=0; j--){
 	  glVertex3f(pipe[i].vertices[j][0], pipe[i].vertices[j][1], -1*pipe[i].vertices[j][2]);
 	}
@@ -108,14 +110,29 @@ void setup_scene(){
 
   // Draw an asteroid
   glPushMatrix();
-  glTranslatef(0, 0, -50-zpos);
-  glRotatef(-1.5*zpos, 0.4, 0.7, 1);
+  glTranslatef((20*r1)-10, (20*r2)-10, -50-zpos);
[…]

### 2018-04-08 Paper.tex
Explain voxelization and normal trees

@@
 The available boolean operations include unions, intersections, and differences; the types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations)\cite{cs_dictionary}.
-This technique provides a simpler way to create three-dimensional shapes without the need to define each individual vertex or face.\\
+This technique provides a simpler way to create three-dimensional shapes without the need to define each individual vertex or face.CSG objects are organized in binary trees, where leaf nodes are the solid primitives and interior nodes are the boolean operations.
+The object is then created recursively from the leaves up, with each boolean operation being replaced by the object which it created\cite{gold}.
 
-Constructive Solid Geometry are organized in binary trees, where leaf nodes are the solid primitives and interior nodes are the boolean operations.
-The object is then created recursively, with each boolean operation being replaced by the object which it created\cite{gold}.
+\begin{figure}[h]
+  \includegraphics[width=16cm]{CSG}
+  \centering
+  \caption{The three types of CSG operations perfomed on a cube and a sphere: (from left) Union, Intersection, Difference}
+\end{figure}
 \subsubsection{Real-Time Rendering and Image-Based CSG}
 One substantial category in CSG is the difference between Real-Time Rendering of CSG objects and Image-Based CSG objects.
+With real-time rendering, the entirety of the object is rendered, while image-basedd techniques instead pass a two-dimensional image of the CSG shape.
 Real-time rendering poses a number of issues, particularly for shapes that are interactively modified during runtime.
@@
 \subsubsection{Interactive Modification}
+In practice, CSG objects are typically rendered with either raycasting or boundary representations, both of which are fairly slow.
+Boundary representations in particular take a significant amount of time to generate, so the final objects are typically treated as static objects and are not modified\cite{advanced_opengl}.
+For instance, a scene where a cylindrical hole is slowly drilled in another shape would need to continually add more polygons to show the growing hole.
+Calculating the boundaries for each iteration of the shape would be inordinately complex.
+A raycasting method would be able to handle this scenario at the same efficiency as any other scenario, but the algorithm itself is still too slow for real-time rendering.
 \subsubsection{Ray Tracing}
@@
 Two such algorithms are the Goldfeather Algorithm and the layered Goldfeather algorithm.
-
 \subsubsection{Goldfeather Algorithm}
@@
 In the layered Goldfeather approach, the biggest detriment to its speed is reading from the stencil buffer to calculate the maximum depth complexity.
-However, with modern graphics hardware the depth complexity can be calculated without reading directly from the stencil buffer by using extensions built in to NVidia and ATI hardware, such as NV\_occlusion\_query\cite{hardware_csg}.
+However, with modern graphics hardware the depth complexity can be calculated without reading directly from the stencil buffer by using extensions built in to NVidia and ATI hardware, such as NV\_occlusion\_query\cite{hardware_csg}.\\
 
+Another performance-increasing method is to normalize the binary tree, which pushes union operations towards the root of the tree and difference/intersection operations towards the leaves\cite{advanced_opengl}.
+Although normalizing the tree can increase the height and number of primitives in the tree, overall the normal form can be processed more quickly.
+In the normalized tree, no intersection or difference node has a union operator in its left subtree, and a single primitive as its right child.
+As union operations, as opposed to difference and intersection, are additive, placing these nodes nearer to the root of the tree saves time, since the unimportant sections are removed earlier in the rendering process.
+To normalize the tree, several set properties are used recursively on the tree; $X-(Y\cup Z)=(X-Y)-Z$ and $X-(Y\cap Z) = (X-Y)\cup(X-Z)$, for example.
+Once the normal tree is created, extraneous nodes can be pruned; intersections and differences where the two objects do not touch are deleted or replaced with the left child, respectively.\\
+
+A more recent technique involves partitioning the scene into voxels instead of pixels using a method similar to Goldfeather's algorithm above, where each voxel of the scene has a flag determining whether it is inside or outside an object\cite{geoinformation}.
+There are two major variants on the voxelization process: voxelizing and combining each primitive, or voxelizing along rays.
+While the first method requires more memory, it is advantageous if the user requires the voxelized version of each subtree.
+Once this preprocessing phase is completed, inclusion tests can be performed in constant time.
+However, the voxelization method is still unsuitable for real-time use.
 \section{Software}
@@
 The shapes are exported as .stl files, which can be read using the reader function located in \texttt{stlReader.cpp}.
-The code within a standard .scad file will look similar to this:
+An example of a typical .scad file is reproduced below.
 \lstset{language=scad}
+\begin{figure}[h]
 \begin{lstlisting}
@@
 \end{lstlisting}
-The \texttt{intersection()} function takes no inputs in its declaration, instead operating on the objects created within its code block.
-\texttt{union()} and \texttt{difference()} work in the same way.
+\centering
+\caption{An exampe code segment from a .scad file}
+\end{figure}
+
+The \texttt{intersection()} function takes no inputs in its declaration, instead operating on the objects created within its code block; \texttt{union()} and \texttt{difference()} work in the same way.
 The \texttt{cube()} function can either create a regular cube when passed a single number, as above, or create rectangular prisms with user-defined dimensions.
-Instead of a single number, the user would pass the dimensions as such: \texttt{cube([15,10,12], center=true);}.
+Instead of a single number, the user would pass the dimensions as such: \texttt{cube([15,10,12]);}.
 The \texttt{center=true} parameter sets the cube with its center at the origin; a parameter of \texttt{center=false} would draw the cube with its bottom corner at the origin.
@@
 In order to manipulate the objects created with OpenSCAD, the CSG objects were exported in .stl files. Each file is formatted in the same way:
+\begin{figure}[h]
 \begin{lstlisting}
@@
 \end{lstlisting}
+\centering
+\caption{A single triangle as saved in an ASCII .stl file}
+\end{figure}
+
 The object is broken down into a series of triangles and a normal vector is calculated.
@@
 In order to solve this issue, the z-coordinate of each vertex was multiplied by -1.
+In conjunction with this fix, the x-value of each obstacle's normal vector was multiplied by -1.
+This change produced better lighting effects on the obstacles with the new vertex values.
+The normal vector for the spaceship was left as is.
 \begin{figure}[h]
@@
 \subsubsection*{die.scad}
+% \lstinputlisting[language=scad]{die.scad}
 \subsubsection*{connector.scad}
+% \lstinputlisting[language=scad]{connector.scad}
 \subsubsection*{torus.scad}
@@
 \subsubsection*{spaceship.scad}
+% \lstinputlisting[language=scad]{spaceship.scad}
 \subsubsection*{cookie\_cutter.scad}
+% \lstinputlisting[language=scad]{cookie_cutter.scad}
 

### 2018-04-08 Paper.tex
Wrote more of the paper

@@
 One substantial category in CSG is the difference between Real-Time Rendering of CSG objects and Image-Based CSG objects.
-With real-time rendering, the entirety of the object is rendered, while image-basedd techniques instead pass a two-dimensional image of the CSG shape.
+With real-time rendering, the entirety of the object is rendered, while image-basedd techniques instead pass a two-dimensional image of the CSG shape from a particular angle.
 Real-time rendering poses a number of issues, particularly for shapes that are interactively modified during runtime.
@@
 Then, the same comparison is made with the second object stored in \texttt{ZTEMP} and compared between the front and back faces of the first object.
-Only the pixels with the flag set for both sets are drawn.
+Only the pixels with the flag set for both sets are drawn.\\
 
+This algorithm does not directly visualize the model; instead, in its transformation to a usable set it is possible that the boolean expression will grow exponentially.
+Thus, the layered version of the Goldfeather algorithm is generally more practical\cite{interactive_csg}.
 \subsubsection{Layered Goldfeather Algorithm}
@@
 Once this preprocessing phase is completed, inclusion tests can be performed in constant time.
-However, the voxelization method is still unsuitable for real-time use.
+However, the voxelization method is still unsuitable for real-time use.\\
+
+Many of the other algorithms that are developed for CSG are either unsuitable for the hardware of graphical processing units (GPUs).
+In recent years, the GPU memory capacity has grown at a slower rate than the computational power of the GPUs themselves, creating a bottleneck for algorithms that rely on the GPU\cite{interactive_csg}.
+For example, raytracing algorithms for CSG divide the rendered object into a set of intervals between when the primitives are intersected, but GPUs cannot store all the intervals for thousands of individual rays unless the scene is relatively simple.
+Another approach relies on recursively raytracing the left and right subtrees, recording where the ray enters, exits, and misses an object.
+Again, this solution is not ideal for a GPU because of the memory needed to store the three states of the ray and the recursive nature of the algorithm\cite{interactive_csg}.
 \section{Software}

### 2018-04-08 game.cpp

@@ -32,9 +32,7 @@ GLfloat light_specular[] = {1, 1, 1, 1.0};
 GLfloat light_position[] = {0.0, 3.0, -5.0, 0.0};
 
 std::vector<Facet> asteroids = readFile("asteroid.stl");
-std::vector<Facet> pipe = readFile("pipes.stl");
 std::vector<Facet> spaceship = readFile("spaceship.stl");
-std::vector<Facet> connector = readFile("connector.stl");
 
 bool upKey, leftKey, rightKey, downKey;
 float r1, r2, r3;
@@ -93,26 +91,11 @@ void setup_scene(){
   glEnd();
   glPopMatrix();
 
-  // Draw a pipe
+  // Draw a series of asteroids
+  glColor3f(0.5, 0.25, 0.25);
   glPushMatrix();
   glTranslatef((20*r3)-10, (20*r1)-10, -25-zpos);
   glRotatef(-1.5*zpos, r1, r2, r3);
-  glColor3f(0.6, 0.6, 0.6);
-  glBegin(GL_TRIANGLES);
-  for(int i = 0; i<pipe.size(); i++){
-	glNormal3f(-1*pipe[i].normal[0], pipe[i].normal[1], pipe[i].normal[2]);
-	for(int j = 2; j>=0; j--){
-	  glVertex3f(pipe[i].vertices[j][0], pipe[i].vertices[j][1], -1*pipe[i].vertices[j][2]);
-	}
-  }
-  glEnd();
-  glPopMatrix();
-
-  // Draw an asteroid
-  glPushMatrix();
-  glTranslatef((20*r1)-10, (20*r2)-10, -50-zpos);
-  glRotatef(-1.5*zpos, -r3, r1, r2);
-  glColor3f(0.5, 0.25, 0.25);
   glBegin(GL_TRIANGLES);
   for(int i = 0; i<asteroids.size(); i++){
   	glNormal3f(-1*asteroids[i].normal[0], asteroids[i].normal[1], asteroids[i].normal[2]);
@@ -123,16 +106,27 @@ void setup_scene(){
   glEnd();
   glPopMatrix();
 
-  // Draw a hollow die
+  glPushMatrix();
+  glTranslatef((20*r1)-10, (20*r2)-10, -50-zpos);
+  glRotatef(-1.5*zpos, -r3, r1, r2);
+  glBegin(GL_TRIANGLES);
+  for(int i = 0; i<asteroids.size(); i++){
[…]

### 2018-04-08 Paper.tex
Wrote conclusion

@@
 \subsubsection{Ray Tracing}
+When using raytracing algorithms for CSG, the boolean operations can be reduced to one-dimensional subproblems\cite{raytrace}.
+For example, a ray passing through two objects along its path would have line segments representing when the ray entered and exited each primitive.
+These line segments can then be subjected to the boolean operations.
+The resulting line segment will be equivalent to a line segment generated from a ray passing through the complete CSG primitive.
+Although individual ray calculations are fast, the excessive number of rays can make more complicated scenes unsuitable for real-time rendering\cite{interactive_csg}.
 \subsection{CSG Algorithms}
@@
 \section{Conclusion}
-
+As a whole, the field of Constructive Solid Geometry poses a number of difficult problems for computer scientists.
+With real-time rendering infeasible, and the alternative algorithms only suitable in a handful of cases, there is still much research to be done.
+However, the existing solutions are more than adequate to produce usable and complete software, as evidenced by the display program and the asteroid game.
 \newpage
@@
 \subsubsection*{torus.scad}
+% \lstinputlisting[language=scad]{torus.scad}
 \subsubsection*{asteroid.scad}
+% \lstinputlisting[language=scad]{asteroid.scad}
 \subsubsection*{pipes.scad}
+% \lstinputlisting[language=scad]{pipes.scad}
 \subsubsection*{spaceship.scad}

### 2018-04-23 Paper.tex
Incorporated edits

@@
 Constructive Solid Geometry (CSG) is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
-The available boolean operations include unions, intersections, and differences; the types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations)\cite{cs_dictionary}.
-This technique provides a simpler way to create three-dimensional shapes without the need to define each individual vertex or face.CSG objects are organized in binary trees, where leaf nodes are the solid primitives and interior nodes are the boolean operations.
+The available boolean operations include unions, intersections, and differences.
+The types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations) \cite{cs_dictionary}.
+This technique provides a simpler way to create three-dimensional shapes without the need to define each individual vertex or face.
+CSG objects are organized in binary trees, where leaf nodes are the solid primitives and interior nodes are the boolean operations.
 The object is then created recursively from the leaves up, with each boolean operation being replaced by the object which it created\cite{gold}.
@@
 Real-time rendering poses a number of issues, particularly for shapes that are interactively modified during runtime.
-While real-time rendering can be accomplished by passing the shape as a mathematically calculated model, this solution is more applicable to static CSG shapes and does not work well for interactive modifications\cite{open_csg}.
-The alternative solution, image-based rendering algorithms, avoid most of the complex boundary calculations of constructive solid geometry by storing only an image of the shape in the frame buffer\cite{open_csg}.
+While real-time rendering can be accomplished by passing the shape as a mathematically calculated model, this solution is more applicable to static CSG shapes and does not work well for interactive modifications \cite{open_csg}.
+The alternative solution, image-based rendering algorithms, avoid most of the complex boundary calculations of constructive solid geometry by storing only an image of the shape in the frame buffer \cite{open_csg}.
 \subsection{Use in Computer Graphics}
@@
 In practice, CSG objects are typically rendered with either raycasting or boundary representations, both of which are fairly slow.
-Boundary representations in particular take a significant amount of time to generate, so the final objects are typically treated as static objects and are not modified\cite{advanced_opengl}.
+Boundary representations in particular take a significant amount of time to generate, so the final objects are typically treated as static objects and are not modified \cite{advanced_opengl}.
 For instance, a scene where a cylindrical hole is slowly drilled in another shape would need to continually add more polygons to show the growing hole.
@@
 \subsubsection{Ray Tracing}
-When using raytracing algorithms for CSG, the boolean operations can be reduced to one-dimensional subproblems\cite{raytrace}.
+When using raytracing algorithms for CSG, the boolean operations can be reduced to one-dimensional subproblems \cite{raytrace}.
 For example, a ray passing through two objects along its path would have line segments representing when the ray entered and exited each primitive.
@@
 The resulting line segment will be equivalent to a line segment generated from a ray passing through the complete CSG primitive.
-Although individual ray calculations are fast, the excessive number of rays can make more complicated scenes unsuitable for real-time rendering\cite{interactive_csg}.
+Although individual ray calculations are fast, the excessive number of rays can make more complicated scenes unsuitable for real-time rendering \cite{interactive_csg}.
 \subsection{CSG Algorithms}
@@
 \subsubsection{Goldfeather Algorithm}
-Goldfeather's algorithm quickly became the precursor to all image-based CSG algorithms\cite{hardware_csg}.
+Goldfeather's algorithm quickly became the precursor to all image-based CSG algorithms \cite{hardware_csg}.
 The algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effectively use the z-buffer on supported graphics hardware.
-The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm, which traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers\cite{gold}.
-Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates.\\
+The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm, which traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers \cite{gold}.
+Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates and $A$ through $F$ are constants.\\
 
@@
 This algorithm does not directly visualize the model; instead, in its transformation to a usable set it is possible that the boolean expression will grow exponentially.
-Thus, the layered version of the Goldfeather algorithm is generally more practical\cite{interactive_csg}.
+Thus, the layered version of the Goldfeather algorithm is generally more practical \cite{interactive_csg}.
 \subsubsection{Layered Goldfeather Algorithm}
@@
 al.
-exploit instances where the maximum depth complexity $k$ (the number of primitives covering each pixel) is less than the number of primitives $n$ in a normalized CSG tree\cite{layered_gold}.
+exploit instances where the maximum depth complexity $k$ (the number of primitives covering each pixel) is less than the number of primitives $n$ in a normalized CSG tree \cite{layered_gold}.
 For example, a scene with five spheres in a line will have no overlapping objects when the viewer is orthogonal to the line.
@@
 Parity is stored as a flag in the algorithm; parity is even when outside an object and odd when inside an object.
-While generally effective, this definition places important constraints on the usable surfaces: the surface cannot self-intersect, and all volumes must be completely covered in a surface with no holes\cite{layered_gold}.\\
+While generally effective, this definition places important constraints on the usable surfaces: the surface cannot self-intersect, and all volumes must be completely covered in a surface with no holes \cite{layered_gold}.\\
 
-The original Goldfeather algorithm performs a single clipping operation in $an+b$ time, where $a$ and $b$ are constants determined by polygon rasterization speed and z-buffer copying, respectively\cite{layered_gold}.
+The original Goldfeather algorithm performs a single clipping operation in $an+b$ time, where $a$ and $b$ are constants determined by polygon rasterization speed and z-buffer copying, respectively \cite{layered_gold}.
 Each primitive is clipped once, giving an upper runtime bound of $O(n^2)$.
@@
 Each layer is copied into the z-buffer, so the time to clip a single layer in the layered approach is $2an+b$.
-In total, with each layer clipped once, the layered algorithm runs in $2akn+bk$, or $O(kn)$\cite{layered_gold}.
+In total, with each layer clipped once, the layered algorithm runs in $2akn+bk$, or $O(kn)$ \cite{layered_gold}.
 As stated above, the algorithm performs better than the original Goldfeather method when $k<n$.
@@
 In OORT, ray tracing calculations between two objects on a CSG tree always test against the left object first.
-Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE\cite{raytrace}.\\
+Thus, functions like \texttt{CSGIntersection} can avoid testing rays against more complex shapes if the left shape returns FALSE \cite{raytrace}.\\
 
@@
 One workaround is to save the second z-buffer into another, unused buffer, such as the color buffer.
-Commonly referred to as render-to-texture, color data is discarded and the texels of the new p-buffer have a one-to-one match to the pixels in the frame buffer\cite{hardware_csg}.\\
+Commonly referred to as render-to-texture, color data is discarded and the texels of the new p-buffer have a one-to-one match to the pixels in the frame buffer \cite{hardware_csg}.\\
 
 In the layered Goldfeather approach, the biggest detriment to its speed is reading from the stencil buffer to calculate the maximum depth complexity.
-However, with modern graphics hardware the depth complexity can be calculated without reading directly from the stencil buffer by using extensions built in to NVidia and ATI hardware, such as NV\_occlusion\_query\cite{hardware_csg}.\\
+However, with modern graphics hardware the depth complexity can be calculated without reading directly from the stencil buffer by using extensions built in to NVidia and ATI hardware, such as NV\_occlusion\_query \cite{hardware_csg}.\\
 
-Another performance-increasing method is to normalize the binary tree, which pushes union operations towards the root of the tree and difference/intersection operations towards the leaves\cite{advanced_opengl}.
+Another performance-increasing method is to normalize the binary tree, which pushes union operations towards the root of the tree and difference/intersection operations towards the leaves \cite{advanced_opengl}.
 Although normalizing the tree can increase the height and number of primitives in the tree, overall the normal form can be processed more quickly.
@@
 
-A more recent technique involves partitioning the scene into voxels instead of pixels using a method similar to Goldfeather's algorithm above, where each voxel of the scene has a flag determining whether it is inside or outside an object\cite{geoinformation}.
+A more recent technique involves partitioning the scene into voxels (small cubes) instead of pixels using a method similar to Goldfeather's algorithm above, where each voxel of the scene has a flag determining whether it is inside or outside an object \cite{geoinformation}.
 There are two major variants on the voxelization process: voxelizing and combining each primitive, or voxelizing along rays.
@@
 Many of the other algorithms that are developed for CSG are either unsuitable for the hardware of graphical processing units (GPUs).
-In recent years, the GPU memory capacity has grown at a slower rate than the computational power of the GPUs themselves, creating a bottleneck for algorithms that rely on the GPU\cite{interactive_csg}.
+In recent years, the GPU memory capacity has grown at a slower rate than the computational power of the GPUs themselves, creating a bottleneck for algorithms that rely on the GPU \cite{interactive_csg}.
 For example, raytracing algorithms for CSG divide the rendered object into a set of intervals between when the primitives are intersected, but GPUs cannot store all the intervals for thousands of individual rays unless the scene is relatively simple.
 Another approach relies on recursively raytracing the left and right subtrees, recording where the ray enters, exits, and misses an object.
-Again, this solution is not ideal for a GPU because of the memory needed to store the three states of the ray and the recursive nature of the algorithm\cite{interactive_csg}.
+Again, this solution is not ideal for a GPU because of the memory needed to store the three states of the ray and the recursive nature of the algorithm \cite{interactive_csg}.
 \section{Software}
@@
 OpenSCAD uses a right-handed coordinate system, while OpenGL uses a left-handed system.
+\begin{figure}[h]
+  \includegraphics[width=12cm]{Right-Hand}
+  \centering
+  \caption{The OpenSCAD viewer.
+As indicated by the legend in the bottom-left corner, the program uses a right-handed coordinate system (the positive z-axis is pointing up).
+In a left-handed coordinate system, such as the one used by OpenGL, the negative z-axis would be pointing up.}
+\end{figure}
 As the objects moved towards and away from the camera, parts of the objects that should have been hidden were visible, and vice versa.
@@
 \section{Conclusion}
-As a whole, the field of Constructive Solid Geometry poses a number of difficult problems for computer scientists.
+As a whole, the field of constructive solid geometry poses a number of difficult problems for computer scientists.
 With real-time rendering infeasible, and the alternative algorithms only suitable in a handful of cases, there is still much research to be done.

### 2018-04-23 Paper.tex
Continued edits

@@
 Constructive Solid Geometry (CSG) is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
-The available boolean operations include unions, intersections, and differences.
+The available boolean operations include union, intersection, and difference.
 The types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations) \cite{cs_dictionary}.
@@
 A raycasting method would be able to handle this scenario at the same efficiency as any other scenario, but the algorithm itself is still too slow for real-time rendering.
-\subsubsection{Ray Tracing}
+\subsubsection{Raytracing}
+In comparison with raycasting, which determines only the first object struck by a light ray, raytracing tracks all objects struck by a light ray, making it far more useful for interactive environments.
 When using raytracing algorithms for CSG, the boolean operations can be reduced to one-dimensional subproblems \cite{raytrace}.
-For example, a ray passing through two objects along its path would have line segments representing when the ray entered and exited each primitive.
+For example, a ray passing through two objects along its path has a line segments representing when the ray entered and exited each primitive.
 These line segments can then be subjected to the boolean operations.
 The resulting line segment will be equivalent to a line segment generated from a ray passing through the complete CSG primitive.
-Although individual ray calculations are fast, the excessive number of rays can make more complicated scenes unsuitable for real-time rendering \cite{interactive_csg}.
+Although individual ray calculations are fast, the excessive number of rays needed to accurately reproduce the model can make more complicated scenes unsuitable for real-time rendering \cite{interactive_csg}.
 \subsection{CSG Algorithms}
 Given the rendering complexity of constructive solid geometry, several algorithms have been developed to improve the speed and quality of CSG primitives.
-Two such algorithms are the Goldfeather Algorithm and the layered Goldfeather algorithm.
+Two such algorithms are the Goldfeather Algorithm and the layered Goldfeather algorithm \cite{hardware_csg}.
 \subsubsection{Goldfeather Algorithm}
 Goldfeather's algorithm quickly became the precursor to all image-based CSG algorithms \cite{hardware_csg}.
-The algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effectively use the z-buffer on supported graphics hardware.
+The algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effective use of the z-buffer on supported graphics hardware.
 The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm, which traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers \cite{gold}.
-Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates and $A$ through $F$ are constants.\\
+Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates and $A$ through $F$ are constants.
+This quadratic equation is used to define the boundary surfaces of all primitive solids in the implementation, including cones, cylinders, and ellipsoids \cite{gold}.\\
 
@@
 Then, the same comparison is made with the second object stored in \texttt{ZTEMP} and compared between the front and back faces of the first object.
-Only the pixels with the flag set for both sets are drawn.\\
-
-This algorithm does not directly visualize the model; instead, in its transformation to a usable set it is possible that the boolean expression will grow exponentially.
-Thus, the layered version of the Goldfeather algorithm is generally more practical \cite{interactive_csg}.
+Only the pixels with the flag set for both sets are drawn.
 \subsubsection{Layered Goldfeather Algorithm}
@@
 exploit instances where the maximum depth complexity $k$ (the number of primitives covering each pixel) is less than the number of primitives $n$ in a normalized CSG tree \cite{layered_gold}.
-For example, a scene with five spheres in a line will have no overlapping objects when the viewer is orthogonal to the line.
+For example, a scene with five spheres in a line has no overlapping objects when the viewer is orthogonal to the line.
 In this case, $k=1$ and is less than the number of primitives, $n=5$.
-If the viewer looked directly down the line, the depth complexity would equal the number of primitives and would not be suitable for the layered algorithm.\\
+If the viewer looked directly down the line, the depth complexity equals the number of primitives and is not suitable for the layered algorithm.\\
 
@@
 The shapes are exported as .stl files, which can be read using the reader function located in \texttt{stlReader.cpp}.
-An example of a typical .scad file is reproduced below.
+An example of a typical .scad file is shown in Figure 2.\\
 \lstset{language=scad}
@@
 The \texttt{intersection()} function takes no inputs in its declaration, instead operating on the objects created within its code block; \texttt{union()} and \texttt{difference()} work in the same way.
-The \texttt{cube()} function can either create a regular cube when passed a single number, as above, or create rectangular prisms with user-defined dimensions.
-Instead of a single number, the user would pass the dimensions as such: \texttt{cube([15,10,12]);}.
-The \texttt{center=true} parameter sets the cube with its center at the origin; a parameter of \texttt{center=false} would draw the cube with its bottom corner at the origin.
+The \texttt{cube()} function can either create a regular cube when passed a single number, as in Figure 2, or create rectangular prisms with user-defined dimensions.
+Instead of a single number, the user  passes the dimensions as such: \texttt{cube([15,10,12]);}.
+The \texttt{center=true} parameter sets the cube with its center at the origin; a parameter of \texttt{center=false} draws the cube with its bottom corner at the origin.
 Cylinders have independently defined radii for both ends; a cone can be created by setting one radius to 0.
 As seen in the \texttt{rotate} call, vectors are contained within square brackets.
-For the sake of simplicity, and to better explore the possibilities of CSG, only rectangular prisms, cones, cylinders, and spheres were used to build each object.
-For transformations, only rotate and translate were used.\\
+For the sake of simplicity, and to better explore the possibilities of CSG, only rectangular prisms, cones, cylinders, and spheres are used to build each object.
+For transformations, only rotate and translate are used.\\
 
-In order to manipulate the objects created with OpenSCAD, the CSG objects were exported in .stl files. Each file is formatted in the same way:
+In order to manipulate the objects created with OpenSCAD, the CSG objects were exported in .stl files. Each file is formatted as shown in Figure 3:
 \begin{figure}[h]
@@
 The object is broken down into a series of triangles and a normal vector is calculated.
-Using this pattern, the .stl files were read by the helper program, stlReader.cpp.
-stlReader.cpp uses the struct \texttt{Facet} to store the data for each triangle, with each \texttt{Facet} being added to a vector containing all the pieces of the shape.
+Using this pattern, the .stl files are read by the helper program, stlReader.cpp.
+stlReader.cpp uses the struct \texttt{Facet} to store the data for each triangle, with each \texttt{Facet} being added to a vector containing all the triangles of the shape.
 The helper program is currently only configured to read .stl files that use the ASCII-style layout, and cannot read binary .stl files.\\
 
-In total, seven objects were created using OpenSCAD. These include, but are not limited to: a die, a torus, an asteroid, a spaceship, and a 6-sided cookie-cutter.
+In total, seven objects are created using OpenSCAD. These include, but are not limited to: a die, a torus, an asteroid, a spaceship, and a 6-sided cookie-cutter.
+\begin{figure}[h]
+  \includegraphics[width=9cm]{Cookie}
+  \centering
+  \caption{A 6-sided cookie cutter as seen in the OpenSCAD environment. The shapes include the four card suits (heart, club, diamond, spade), a cross, and a star.}
+\end{figure}
 \subsubsection{OpenGL Implementation}
@@
 Within the \texttt{glBegin} and \texttt{glEnd} block, the vector \texttt{newShape} is scanned, with each \texttt{Facet}'s normal vector and vertices passed to the respective OpenGL command.
-In order to correctly light the triangle elements, the vertices are passed in the reverse order of how were printed in the .stl file.
+In order to correctly light the triangle elements, the vertices are passed in the reverse order of how they are printed in the .stl file.
 \subsection{Asteroid Game}
+\begin{figure}
+  \includegraphics[width=11cm]{Game}
+  \centering
+  \caption{A screenshot from the asteroid game.}
+\end{figure}
 The second implementation of constructive solid geometry is an interactive game, where the user controls a spaceship and avoids asteroids.
@@
 When the scene is redrawn, the camera position uses \texttt{xpos} and \texttt{ypos} to remain behind the ship.
-Another variable, \texttt{zpos} is also changed; this variable does not tranlate the spaceship itself, but instead translates all other obstacles onscreen.
+Another variable, \texttt{zpos} is also changed; this variable does not translate the spaceship itself, but instead translates all other obstacles onscreen.
 The \texttt{zpos} variable is cyclical; once the value is less than -100, \texttt{zpos} is reset to 0, and the \texttt{speed}, \texttt{zpos}'s rate of change, is increased.\\
@@
 For example, the spaceship engines jut out from the back of the ship.
-Thus, a viewer positioned in between the two engines should be able to see the right curve of the left engine and the left curve of the right engine, as shown below left.
+Thus, a viewer positioned in between the two engines should be able to see the right curve of the left engine and the left curve of the right engine, as shown in Figure 4.
 However, the viewer instead could see the left curve of the left engine and the right curve of the right engine, as seen below right; this perspective made the engines appear recessed instead of external.

### 2018-04-24 game.cpp
Add more asteroids and fix random generator

@@ -25,8 +25,8 @@ GLint window_width = 700;
 GLint window_height = 700;
 static GLfloat xpos = 0;
 static GLfloat ypos = 0;
-static GLfloat zpos = 0;
-static GLfloat speed = 0.2;
+static GLfloat zpos = 25;
+static GLfloat speed = 0.5;
 
 GLfloat light_specular[] = {1, 1, 1, 1.0};
 GLfloat light_position[] = {0.0, 3.0, -5.0, 0.0};
@@ -34,18 +34,31 @@ GLfloat light_position[] = {0.0, 3.0, -5.0, 0.0};
 std::vector<Facet> asteroids = readFile("asteroid.stl");
 std::vector<Facet> spaceship = readFile("spaceship.stl");
 
+struct GameObject{
+  double position[3];
+  double rotation[3];
+};
+
+std::vector<GameObject> obstacles;
+
 bool upKey, leftKey, rightKey, downKey;
-float r1, r2, r3;
 
 void flight(){
   // Change spaceship position
   zpos -= speed;
   if(zpos <= -100){
-	r1 = float(rand())/RAND_MAX;
-	r2 = float(rand())/RAND_MAX;
-	r3 = float(rand())/RAND_MAX;
-	zpos = 0;
+	zpos = 25;
 	speed += 0.01;
+	if(speed*10 == int(speed*10)){
+	  GameObject tmp;
+	  obstacles.push_back(tmp);
+	}
+	for(int i=0; i<obstacles.size(); i++){
+	  for(int j=0; j<3; j++){
+		obstacles[i].position[j]=float(rand())/RAND_MAX;
+		obstacles[i].rotation[j]=float(rand())/RAND_MAX;
+	  }
+	}
   }
   if(leftKey){
 	if(xpos > -10) xpos -= 0.2;
@@ -91,46 +104,61 @@ void setup_scene(){
   glEnd();
   glPopMatrix();
 
-  // Draw a series of asteroids
   glColor3f(0.5, 0.25, 0.25);
-  glPushMatrix();
-  glTranslatef((20*r3)-10, (20*r1)-10, -25-zpos);
-  glRotatef(-1.5*zpos, r1, r2, r3);
-  glBegin(GL_TRIANGLES);
-  for(int i = 0; i<asteroids.size(); i++){
[…]

### 2018-04-24 Paper.tex
Incorporate more edits

@@
 \begin{document}
+\begin{titlepage}
 \title{Constructive Solid Geometry for Interactive Environments}
@@
 \maketitle
+\begin{abstract}
+  One barrier to high-quality computer generated images is the exorbitant number of polygons and points needed to define a detailed object.
+Defining each vertex by hand would take a user hours for even the simplest shapes.
+One way to streamline the process is to use Constructive Solid Geometry (CSG) to define shapes as combinations of geometric primitives.
+This paper outlines the advantages and disadvantages of this method, as well as two major algorithms used to render CSG images.
+Furthermore, an implementation of CSG shapes using OpenGL is described.
+Although adequate for the purpose of th ecreated software, there is still room for improvement in CSG usage and implementation.
+\end{abstract}
+\end{titlepage}
+\begin{onehalfspace}
+\tableofcontents
+\listoffigures
+\newpage
+\end{onehalfspace}
 \begin{doublespace}
@@
 \subsection{General Concepts}
-Constructive Solid Geometry (CSG) is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
+Constructive solid geometry is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
 The available boolean operations include union, intersection, and difference.
@@
 For example, the spaceship engines jut out from the back of the ship.
-Thus, a viewer positioned in between the two engines should be able to see the right curve of the left engine and the left curve of the right engine, as shown in Figure 4.
+Thus, a viewer positioned in between the two engines should be able to see the right curve of the left engine and the left curve of the right engine, as shown in Figure 7.
 However, the viewer instead could see the left curve of the left engine and the right curve of the right engine, as seen below right; this perspective made the engines appear recessed instead of external.
@@
 \begin{onehalfspace}
-\section*{Source Code}
-\subsection*{OpenSCAD}
-\subsubsection*{die.scad}
+\section{Source Code}
+\subsection{OpenSCAD}
+\subsubsection{die.scad}
 % \lstinputlisting[language=scad]{die.scad}
-\subsubsection*{connector.scad}
-% \lstinputlisting[language=scad]{connector.scad}
-\subsubsection*{torus.scad}
-% \lstinputlisting[language=scad]{torus.scad}
-\subsubsection*{asteroid.scad}
-% \lstinputlisting[language=scad]{asteroid.scad}
-\subsubsection*{pipes.scad}
-% \lstinputlisting[language=scad]{pipes.scad}
-\subsubsection*{spaceship.scad}
-% \lstinputlisting[language=scad]{spaceship.scad}
-\subsubsection*{cookie\_cutter.scad}
-% \lstinputlisting[language=scad]{cookie_cutter.scad}
+\subsubsection{connector.scad}
+%\lstinputlisting[language=scad]{connector.scad}
+\subsubsection{torus.scad}
+\lstinputlisting[language=scad]{torus.scad}
+\subsubsection{asteroid.scad}
+\lstinputlisting[language=scad]{asteroid.scad}
+\subsubsection{pipes.scad}
+\lstinputlisting[language=scad]{pipes.scad}
+\subsubsection{spaceship.scad}
+\lstinputlisting[language=scad]{spaceship.scad}
+\subsubsection{cookie\_cutter.scad}
+\lstinputlisting[language=scad]{cookie_cutter.scad}
 
-\subsection*{C++}
-\subsubsection*{stlReader.cpp}
-% \lstinputlisting[language=C++]{stlReader.cpp}
-\subsubsection*{display.cpp}
-% \lstinputlisting[language=C++]{display.cpp}
-\subsubsection*{game.cpp}
-% \lstinputlisting[language=C++]{game.cpp}
+\subsection{C++}
+\subsubsection{stlReader.cpp}
+\lstinputlisting[language=C++]{stlReader.cpp}
+\subsubsection{display.cpp}
+\lstinputlisting[language=C++]{display.cpp}
+\subsubsection{game.cpp}
+\lstinputlisting[language=C++]{game.cpp}
 

### 2018-04-24 game.cpp

@@ -48,11 +48,7 @@ void flight(){
   zpos -= speed;
   if(zpos <= -100){
 	zpos = 25;
-	speed += 0.01;
-	if(speed*10 == int(speed*10)){
-	  GameObject tmp;
-	  obstacles.push_back(tmp);
-	}
+	if(speed<1.2) speed += 0.02;
 	for(int i=0; i<obstacles.size(); i++){
 	  for(int j=0; j<3; j++){
 		obstacles[i].position[j]=float(rand())/RAND_MAX;
@@ -107,7 +103,7 @@ void setup_scene(){
   glColor3f(0.5, 0.25, 0.25);
   for(int i=0; i<obstacles.size(); i++){
 	glPushMatrix();
-	glTranslatef(20*obstacles[i].position[0]-10, 20*obstacles[i].position[1]-10, -90*obstacles[i].position[2]-zpos);
+	glTranslatef(30*obstacles[i].position[0]-15, 30*obstacles[i].position[1]-15, -95*obstacles[i].position[2]-zpos);
 	glRotatef(-1.5*zpos, 2*obstacles[i].rotation[0]-1, 2*obstacles[i].rotation[1]-1, 2*obstacles[i].rotation[2]-1);
 	glBegin(GL_TRIANGLES);
 	for(int j=0; j<asteroids.size(); j++){
@@ -119,46 +115,6 @@ void setup_scene(){
 	glEnd();
 	glPopMatrix();
   }
-  // // Draw a series of asteroids
-  // glColor3f(0.5, 0.25, 0.25);
-  // glPushMatrix();
-  // glTranslatef((20*r3)-10, (20*r1)-10, -25-zpos);
-  // glRotatef(-1.5*zpos, r1, r2, r3);
-  // glBegin(GL_TRIANGLES);
-  // for(int i = 0; i<asteroids.size(); i++){
-  // 	glNormal3f(-1*asteroids[i].normal[0], asteroids[i].normal[1], asteroids[i].normal[2]);
-  // 	for(int j = 2; j>=0; j--){
-  // 	  glVertex3f(asteroids[i].vertices[j][0], asteroids[i].vertices[j][1], -1*asteroids[i].vertices[j][2]);
-  // 	}
-  // }
-  // glEnd();
-  // glPopMatrix();
-
-  // glPushMatrix();
-  // glTranslatef((20*r1)-10, (20*r2)-10, -50-zpos);
[…]

### 2018-04-25 game.cpp
Asteroids are added as the game progresses

@@ -25,11 +25,11 @@ GLint window_width = 700;
 GLint window_height = 700;
 static GLfloat xpos = 0;
 static GLfloat ypos = 0;
-static GLfloat zpos = 25;
-static GLfloat speed = 0.5;
+static GLfloat zpos = 50;
+static GLfloat speed = 0.2;
 
 GLfloat light_specular[] = {1, 1, 1, 1.0};
-GLfloat light_position[] = {0.0, 3.0, -5.0, 0.0};
+GLfloat light_position[] = {0.0, 1.0, -3.0, 0.0};
 
 std::vector<Facet> asteroids = readFile("asteroid.stl");
 std::vector<Facet> spaceship = readFile("spaceship.stl");
@@ -37,6 +37,13 @@ std::vector<Facet> spaceship = readFile("spaceship.stl");
 struct GameObject{
   double position[3];
   double rotation[3];
+
+  GameObject(){
+	for(int i=0;i<3;i++){
+	  position[i]=0;
+	  rotation[i]=0;
+	}
+  }
 };
 
 std::vector<GameObject> obstacles;
@@ -47,8 +54,9 @@ void flight(){
   // Change spaceship position
   zpos -= speed;
   if(zpos <= -100){
-	zpos = 25;
-	if(speed<1.2) speed += 0.02;
+	if(speed<1) speed += 0.1;
+	zpos = 50*speed;
+	if(speed*10==int(speed*10)) obstacles.push_back(GameObject());
 	for(int i=0; i<obstacles.size(); i++){
 	  for(int j=0; j<3; j++){
 		obstacles[i].position[j]=float(rand())/RAND_MAX;
@@ -78,6 +86,22 @@ void flight(){
 void setup_scene(){
   glTranslatef(0,0,-10);
   gluLookAt(xpos, ypos+1, 5,    xpos, ypos, 0,     0,1,0);
+  
+  glColor3f(0.5, 0.25, 0.25);
+  for(int i=0; i<obstacles.size(); i++){
+	glPushMatrix();
+	glTranslatef(30*obstacles[i].position[0]-15, 30*obstacles[i].position[1]-15, -90*obstacles[i].position[2]-zpos);
+	glRotatef(-1.5*zpos, 2*obstacles[i].rotation[0]-1, 2*obstacles[i].rotation[1]-1, 2*obstacles[i].rotation[2]-1);
[…]

### 2018-04-26 Paper.tex
Update paper with new random functions

@@
 Another variable, \texttt{zpos} is also changed; this variable does not translate the spaceship itself, but instead translates all other obstacles onscreen.
-The \texttt{zpos} variable is cyclical; once the value is less than -100, \texttt{zpos} is reset to 0, and the \texttt{speed}, \texttt{zpos}'s rate of change, is increased.\\
+The \texttt{zpos} variable is cyclical; once the value is less than -100, \texttt{zpos} is reset to The ``playing area'' of the game ends when \texttt{zpos=-100}, at which point it resets to \texttt{zpos = 50*speed+25}, and the \texttt{speed}, \texttt{zpos}'s rate of change, is increased by 0.2 to a max speed of 1.\\
 
@@
 
-Three random floating-point variables between 0 and 1 are created on each cycle; these three variables determine the rotation and position of each obstacle onscreen.
-For the \texttt{glTranslate} calls, the random numbers are multiplied by 20 then subtracted by 10 to place it within the playing area; the \texttt{glRotate} calls use the random values as they are.
-Each of the three variables is shuffled in their usage in the objects, so that no object has the same rotation or position as another.
+Another struct, \texttt{GameObject}s, is used to hold two three-element arrays; these three variables determine the rotation and position of each obstacle onscreen.
+A vector of these \texttt{GameObject}s is created in the \texttt{init} function and populated with 7 asteroids.
+In a for loop, each obstacle in the vector is drawn with its corresponding rotation and position.
+For the \texttt{glTranslate} calls, the following equation is used to center the asteroids in the xy-plane of the game screen: \texttt{30*obstacles[i].position[0]-15} for x and \texttt{30*obstacles[i].position[1]-15}for y.
+The bounds on the spaceship are between -10 and 10 on both the x- and y-axis, and this formula bounds the position of the asteroid between -15 and 15.
+The z-position of the asteroid is determined by \texttt{-90*obstacles[i].position[2]-zpos}.
+\\
+
+The \texttt{glRotate} calls use the random values with the equation \texttt{2*obstacles[i].rotation[0]-1} for x, \texttt{2*obstacles[i].rotation[1]-1}, for y, and \texttt{2*obstacles[i].rotation[2]-1} for the z-value of the rotation vector.
+The angle of rotation is calculated by \texttt{-1.5*zpos}, causing the asteroid to continually rotate as it moves through the scene.
+By multiplying by 2 and subtracting 1, each component of the rotation vector is bounded between -1 and 1.
 \section{Conclusion}

### 2018-04-26 game.cpp

@@ -54,8 +54,8 @@ void flight(){
   // Change spaceship position
   zpos -= speed;
   if(zpos <= -100){
-	if(speed<1) speed += 0.1;
-	zpos = 50*speed;
+	if(speed<1) speed += 0.02;
+	zpos = 50*speed+25;
 	if(speed*10==int(speed*10)) obstacles.push_back(GameObject());
 	for(int i=0; i<obstacles.size(); i++){
 	  for(int j=0; j<3; j++){

### 2018-04-29 Paper.tex
Recreate die and connector OpenSCAD files
Create presentation

@@
 In order to solve this issue, the z-coordinate of each vertex was multiplied by -1.
-In conjunction with this fix, the x-value of each obstacle's normal vector was multiplied by -1.
+In conjunction with this fix, the x- and y-values of each obstacle's normal vector was multiplied by -1.
 This change produced better lighting effects on the obstacles with the new vertex values.
@@
 The bounds on the spaceship are between -10 and 10 on both the x- and y-axis, and this formula bounds the position of the asteroid between -15 and 15.
-The z-position of the asteroid is determined by \texttt{-90*obstacles[i].position[2]-zpos}.
-\\
+The z-position of the asteroid is determined by \texttt{-90*obstacles[i].position[2]-zpos}.\\
 
@@
 With real-time rendering infeasible, and the alternative algorithms only suitable in a handful of cases, there is still much research to be done.
-However, the existing solutions are more than adequate to produce usable and complete software, as evidenced by the display program and the asteroid game.
+Although this OpenGL implementation was not ideal for rendering the more complicated CSG shapes, such as the torus, this solution was more than adequate to produce usable and complete software.
+Future work can be done to increase the speed of the software for these complex shapes, as well as combining CSG and collision detection to improve the asteroid game.
 \newpage
@@
 \subsubsection{die.scad}
-% \lstinputlisting[language=scad]{die.scad}
+\lstinputlisting[language=scad]{die.scad}
 \subsubsection{connector.scad}
-%\lstinputlisting[language=scad]{connector.scad}
+\lstinputlisting[language=scad]{connector.scad}
 \subsubsection{torus.scad}

### 2018-04-29 Paper.tex
Finish edits

@@
 \maketitle
+\end{titlepage}
+\begin{onehalfspace}
+\tableofcontents
+\listoffigures
 \begin{abstract}
@@
 Furthermore, an implementation of CSG shapes using OpenGL is described.
-Although adequate for the purpose of th ecreated software, there is still room for improvement in CSG usage and implementation.
+Although adequate for the purpose of the created software, there is still room for improvement in CSG usage and implementation.
 \end{abstract}
-\end{titlepage}
-\begin{onehalfspace}
-\tableofcontents
-\listoffigures
 \newpage
@@
 \subsection{General Concepts}
-Constructive solid geometry is a method for creating more complicated shapes by performing boolean operations on simpler shapes.
-The available boolean operations include union, intersection, and difference.
-The types of shapes vary from one implementation to another, but generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations) \cite{cs_dictionary}.
+Constructive solid geometry is a method for creating more complicated shapes by performing Boolean operations on simpler shapes.
+The available Boolean operations include union, intersection, and difference.
+The types of shapes vary from one implementation to another, but primitives generally include spheres, cubes, and cylinders, as well as their standard transformations (rotations, scaling, translations) \cite{cs_dictionary}.
 This technique provides a simpler way to create three-dimensional shapes without the need to define each individual vertex or face.
-CSG objects are organized in binary trees, where leaf nodes are the solid primitives and interior nodes are the boolean operations.
-The object is then created recursively from the leaves up, with each boolean operation being replaced by the object which it created\cite{gold}.
+CSG objects are organized in binary trees, where leaf nodes are the solid primitives and interior nodes are the Boolean operations.
+The object is then created recursively from the leaves up, with each Boolean operation being replaced by the object which it created\cite{gold}.
 
@@
   \centering
-  \caption{The three types of CSG operations perfomed on a cube and a sphere: (from left) Union, Intersection, Difference}
+  \caption{The three types of CSG operations performed on a cube and a sphere: (from left) Union, Intersection, Difference}
 \end{figure}
-\subsubsection{Real-Time Rendering and Image-Based CSG}
+\subsection{Real-World Applications}
+CSG is commonly used in a number of computer graphics applications.
+The first version of the Unreal game engine provides a level editor, which included support for real-time CSG operations as early as 1996 \cite{unreal}.
+CSG is also used in raytracing software, such as POV-Ray and Pixar's RenderMan \cite{open_csg}.
+It is also used in a number of computer-assisted design programs, such as OpenSCAD (which was used to create the objects used in this paper).
+\subsection{Real-Time Rendering and Image-Based CSG}
 One substantial category in CSG is the difference between Real-Time Rendering of CSG objects and Image-Based CSG objects.
-With real-time rendering, the entirety of the object is rendered, while image-basedd techniques instead pass a two-dimensional image of the CSG shape from a particular angle.
+With real-time rendering, the entirety of the object is rendered, while image-based techniques instead pass a two-dimensional image of the CSG shape from a particular angle.
 Real-time rendering poses a number of issues, particularly for shapes that are interactively modified during runtime.
-While real-time rendering can be accomplished by passing the shape as a mathematically calculated model, this solution is more applicable to static CSG shapes and does not work well for interactive modifications \cite{open_csg}.
+While real-time rendering can be accomplished by calculating the boundary of the shape and storing it as a polygonal model, this solution is more applicable to static CSG shapes and does not work well if the Boolean CSG operations are modified interactively \cite{open_csg}.
 The alternative solution, image-based rendering algorithms, avoid most of the complex boundary calculations of constructive solid geometry by storing only an image of the shape in the frame buffer \cite{open_csg}.
@@
 \subsubsection{Interactive Modification}
-In practice, CSG objects are typically rendered with either raycasting or boundary representations, both of which are fairly slow.
+In practice, CSG objects are typically rendered with either raycasting (determining the first object struck by a ray) or boundary representations, both of which are fairly slow.
 Boundary representations in particular take a significant amount of time to generate, so the final objects are typically treated as static objects and are not modified \cite{advanced_opengl}.
@@
 In comparison with raycasting, which determines only the first object struck by a light ray, raytracing tracks all objects struck by a light ray, making it far more useful for interactive environments.
-When using raytracing algorithms for CSG, the boolean operations can be reduced to one-dimensional subproblems \cite{raytrace}.
+When using raytracing algorithms for CSG, the Boolean operations can be reduced to one-dimensional subproblems \cite{raytrace}.
 For example, a ray passing through two objects along its path has a line segments representing when the ray entered and exited each primitive.
-These line segments can then be subjected to the boolean operations.
+These line segments can then be subjected to the Boolean operations.
 The resulting line segment will be equivalent to a line segment generated from a ray passing through the complete CSG primitive.
@@
 Given the rendering complexity of constructive solid geometry, several algorithms have been developed to improve the speed and quality of CSG primitives.
-Two such algorithms are the Goldfeather Algorithm and the layered Goldfeather algorithm \cite{hardware_csg}.
+Two such algorithms are the Goldfeather Algorithm \cite{gold} and the layered Goldfeather algorithm \cite{layered_gold}.
 \subsubsection{Goldfeather Algorithm}
@@
 The algorithm, originally designed for the Pixel-Planes graphics hardware, uses partial products to allow effective use of the z-buffer on supported graphics hardware.
-The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm, which traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers \cite{gold}.
+The algorithm is of two parts: the passed CSG tree is first restructured into an equivalent tree more easily used by the second portion of the algorithm.
+The second part traverses the new tree and returns quadratic coefficients to be read by Pixel-Powers \cite{gold}.
 Pixel-Powers, an extension of Pixel-Planes, calculates the values of a quadratic function $Ax^2 + Bxy + Cy^2 + Dx + Ey + F$, where $x$ and $y$ are screen coordinates and $A$ through $F$ are constants.
-This quadratic equation is used to define the boundary surfaces of all primitive solids in the implementation, including cones, cylinders, and ellipsoids \cite{gold}.\\
+This quadratic equation is sufficient to define the boundary surfaces of all primitive solids in the implementation, including cones, cylinders, and ellipsoids \cite{gold}.\\
 
-When perfoming CSG unions in this algorithm, the visible section of the first object is stored in a z-buffer, denoted in Goldfeather's paper as \texttt{ZMIN}, and the visible section of the second object is stored in another z-buffer, denoted \texttt{ZTEMP}.
+When performing CSG unions in this algorithm, the visible section of the first object is stored in a two-dimensional z-buffer, denoted in Goldfeather's paper as \texttt{ZMIN}, and the visible section of the second object is stored in another z-buffer, denoted \texttt{ZTEMP}.
 For all the pixels in the visible area of the second object, a flag \texttt{F1} is set to true.
@@
 al.
-exploit instances where the maximum depth complexity $k$ (the number of primitives covering each pixel) is less than the number of primitives $n$ in a normalized CSG tree \cite{layered_gold}.
+exploit instances where the maximum depth complexity $k$ (the number of primitive solids covering each pixel) is less than the number of primitives $n$ in a normalized CSG tree \cite{layered_gold}.
 For example, a scene with five spheres in a line has no overlapping objects when the viewer is orthogonal to the line.
@@
 The layered algorithm defines layers by their surface parity, or the number of edges in the z-buffer.
-Parity is stored as a flag in the algorithm; parity is even when outside an object and odd when inside an object.
-While generally effective, this definition places important constraints on the usable surfaces: the surface cannot self-intersect, and all volumes must be completely covered in a surface with no holes \cite{layered_gold}.\\
+This is stored as a flag in the algorithm; parity is even when outside an object and odd when inside an object.
+While generally effective, this definition places important constraints on the usable surfaces: the surface cannot self-intersect, and all volumes must be completely covered in a surface with no holes \cite{layered_gold}.
+Otherwise, it would be possible for a ray to become odd when entering an object, but by leaving the object through a hole no intersection is registered and parity does not change.\\
 
-The original Goldfeather algorithm performs a single clipping operation in $an+b$ time, where $a$ and $b$ are constants determined by polygon rasterization speed and z-buffer copying, respectively \cite{layered_gold}.
+The original Goldfeather algorithm performs a single clipping operation in $an+b$ time, where $a$ and $b$ are constants determined by polygon rasterization speed and z-buffer copying, respectively, and there are $n$ primitives in the scene \cite{layered_gold}.
 Each primitive is clipped once, giving an upper runtime bound of $O(n^2)$.
@@
 In addition to the core algorithms, there are other, more minute techniques that can improve performance when rendering CSG objects.
-In Nicholas Wilt's implementation of CSG in the Object Oriented Ray Tracing (OORT) class library, complicated calculations can be short-circuited by creating optimized CSG trees where simpler objects are the left children of each boolean operation.
+In Nicholas Wilt's implementation of CSG in the Object Oriented Ray Tracing (OORT) class library, complicated calculations can be short-circuited by creating optimized CSG trees where simpler objects are the left children of each Boolean operation.
 In OORT, ray tracing calculations between two objects on a CSG tree always test against the left object first.
@@
 
-The Goldfeather algorithm uses two separate z-buffers, which can create issues on systems that do not support multiple z-buffers.
-Alternative approaches, such as copying one set of z-values into main memory, are inadequately slow and values can lose exactness during copying.
+The Goldfeather algorithm uses two separate z-buffers, \texttt{ZMIN} and \texttt{ZTEMP}, which can create issues on systems that do not support multiple z-buffers.
+Alternative approaches, such as copying one set of z-values into main memory, are inadequately slow.
 One workaround is to save the second z-buffer into another, unused buffer, such as the color buffer.
-Commonly referred to as render-to-texture, color data is discarded and the texels of the new p-buffer have a one-to-one match to the pixels in the frame buffer \cite{hardware_csg}.\\
+Commonly referred to as render-to-texture, color data is discarded and the texels of the new buffer have a one-to-one match to the pixels in the frame buffer \cite{hardware_csg}.\\
 
 In the layered Goldfeather approach, the biggest detriment to its speed is reading from the stencil buffer to calculate the maximum depth complexity.
-However, with modern graphics hardware the depth complexity can be calculated without reading directly from the stencil buffer by using extensions built in to NVidia and ATI hardware, such as NV\_occlusion\_query \cite{hardware_csg}.\\
+However, with modern graphics hardware the depth complexity can be calculated without reading directly from the stencil buffer by using extensions built in to Nvidia and ATI hardware, such as NV\_occlusion\_query \cite{hardware_csg}.\\
 
@@
 To normalize the tree, several set properties are used recursively on the tree; $X-(Y\cup Z)=(X-Y)-Z$ and $X-(Y\cap Z) = (X-Y)\cup(X-Z)$, for example.
-Once the normal tree is created, extraneous nodes can be pruned; intersections and differences where the two objects do not touch are deleted or replaced with the left child, respectively.\\
+Once the normal tree is created, extraneous nodes (such as objects intersecting with themselves or objects that are do not intersect at all) can be pruned; intersections and differences where the two objects do not touch are deleted or replaced with the left child, respectively.\\
 
-A more recent technique involves partitioning the scene into voxels (small cubes) instead of pixels using a method similar to Goldfeather's algorithm above, where each voxel of the scene has a flag determining whether it is inside or outside an object \cite{geoinformation}.
+A more recent technique involves partitioning the scene into voxels (the three-dimensional equivalent of a pixel) using a method similar to Goldfeather's algorithm above, where each voxel of the scene has a flag determining whether it is inside or outside an object \cite{geoinformation}.
 There are two major variants on the voxelization process: voxelizing and combining each primitive, or voxelizing along rays.
@@
 However, the voxelization method is still unsuitable for real-time use.\\
+
+\begin{figure}[h]
+  \includegraphics[width=11cm]{Voxel}
+  \centering
+  \caption{A voxelized version of a CSG tree. If needed, the result of the right subtree could be saved for further use}
+\end{figure}
 
@@
 The shapes are exported as .stl files, which can be read using the reader function located in \texttt{stlReader.cpp}.
-An example of a typical .scad file is shown in Figure 2.\\
+An example of a typical .scad file is shown in Figure 3.\\
 \lstset{language=scad}
@@
 \centering
-\caption{An exampe code segment from a .scad file}
+\caption{An example code segment from a .scad file}
 \end{figure}
@@
 The \texttt{intersection()} function takes no inputs in its declaration, instead operating on the objects created within its code block; \texttt{union()} and \texttt{difference()} work in the same way.
-The \texttt{cube()} function can either create a regular cube when passed a single number, as in Figure 2, or create rectangular prisms with user-defined dimensions.
+The \texttt{cube()} function can either create a regular cube when passed a single number, as in Figure 3, or create rectangular prisms with user-defined dimensions.
 Instead of a single number, the user  passes the dimensions as such: \texttt{cube([15,10,12]);}.
@@
 
-In order to manipulate the objects created with OpenSCAD, the CSG objects were exported in .stl files. Each file is formatted as shown in Figure 3:
+In order to manipulate the objects created with OpenSCAD, the CSG objects were exported in .stl files. Each file is formatted as shown in Figure 4.\\
 \begin{figure}[h]
@@
 The helper program is currently only configured to read .stl files that use the ASCII-style layout, and cannot read binary .stl files.\\
+\lstset{language=C++}
+\begin{figure}[h]
+  \begin{lstlisting}
+  struct Facet{
+    double normal[3]; // The x,y,z values of the normal vector
+    double vertices[3][3]; // The x,y,z values of the three vertices for each Facet. [][i] covers the x,y,z values, while [j][] handles the points
+  };
+\end{lstlisting}
+\centering
+\caption{The Facet struct from stlReader.cpp.}
+\end{figure}
 
-In total, seven objects are created using OpenSCAD. These include, but are not limited to: a die, a torus, an asteroid, a spaceship, and a 6-sided cookie-cutter.
+In total, seven objects are created using OpenSCAD. These include: a die, a die with holes through it, a pipe, a torus, an asteroid, a spaceship, and a 6-sided cookie-cutter.
 \begin{figure}[h]
@@
 In order to correctly light the triangle elements, the vertices are passed in the reverse order of how they are printed in the .stl file.
+OpenGL defines front faces by passing vertices in counterclockwise order, whist OpenSCAD lists the vertices in clockwise order.
 \subsection{Asteroid Game}
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=11cm]{Game}
@@
 In comparison to the display program, this program enables the \texttt{GL{\bf\_}COLOR{\bf\_}MATERIAL} state in its \texttt{init} call, which allows the \texttt{glColor} function to be used in conjunction with lighting.
-The spaceship surface is colored a light grey, with the engine trails colored blue; the asteroids are colored red.
+The spaceship surface is colored a light gray, with the engine trails colored blue; the asteroids are colored red.
 For the engine color, the code checks to see if the normal vector is pointing in the same direction as the engines, as well as if the z-value of the triangle's first vertex is at -1.5, at the back of the ship.
@@
 Additionally, more key bindings are added to control the movement of the spaceship.
-Each of the WASD keys has a respective boolean variable that is set to true when the key is pressed and false when the key is released.
+Each of the WASD keys has a respective Boolean variable that is set to true when the key is pressed and false when the key is released.
 This method allows for diagonal movement as well as cardinal movement.
@@
 In the idle loop of the game, the position values of the spaceship are changed.
-If one of the directional booleans is true, \texttt{xpos} or \texttt{ypos} is changed accordingly.
+If one of the directional Booleans is true, \texttt{xpos} or \texttt{ypos} is changed accordingly.
 When the scene is redrawn, the camera position uses \texttt{xpos} and \texttt{ypos} to remain behind the ship.
@@
 The increased movement of the game objects highlighted one major challenge of converting the OpenSCAD designs to OpenGL.
-OpenSCAD uses a right-handed coordinate system, while OpenGL uses a left-handed system.
+OpenSCAD uses a right-handed coordinate system, while OpenGL uses a left-handed system.\\
 \begin{figure}[h]
@@
 For example, the spaceship engines jut out from the back of the ship.
-Thus, a viewer positioned in between the two engines should be able to see the right curve of the left engine and the left curve of the right engine, as shown in Figure 7.
+Thus, a viewer positioned in between the two engines should be able to see the right curve of the left engine and the left curve of the right engine, as shown in Figure 9.
 However, the viewer instead could see the left curve of the left engine and the right curve of the right engine, as seen below right; this perspective made the engines appear recessed instead of external.
@@
 As a whole, the field of constructive solid geometry poses a number of difficult problems for computer scientists.
-With real-time rendering infeasible, and the alternative algorithms only suitable in a handful of cases, there is still much research to be done.
+With real-time rendering unfeasible, and the alternative algorithms only suitable in a handful of cases, there is still much research to be done.
 Although this OpenGL implementation was not ideal for rendering the more complicated CSG shapes, such as the torus, this solution was more than adequate to produce usable and complete software.
@@
 258-267.
+
+\bibitem{unreal}
+  Wilson, Billy, Chris O'Brien, Manveer Heir, and Robert Howarth.
+``Talkin' nasty with Epic's Tim Sweeny.'' {\it VoodoExtreme}, 25 May 2000, web.archive.org/web/20000620093601/http://www.voodooextreme.com/games/interviews/\\timsweeney.html
+
 \end{thebibliography}
